### 2018-04-30 presentation/main.tex
Major commmit
Lot's of stuff here.

\documentclass[draft]{beamer}

\usetheme{Pittsburgh}
\usecolortheme{dove}

\usepackage[]{amsmath}
\usepackage{amssymb}
\usepackage{cmbright}
\usepackage{bm}

\newcommand{\RR}{\mathbb{R}}
\newcommand{\abs}[1]{|#1|}
\newcommand{\set}[1]{\{ #1 \}}
\newcommand{\defkey}{\textbf}

\title{A Stochastic Quasi-Newton Optimizer for \texttt{TensorFlow}}
\author[Chen, Kraemer]{%
Jason Chen\inst{1}, %
David Kraemer\inst{1}}

\institute[Stony Brook University]
{
  \inst{1}%
  CSE 592: Convex Optimization \\
  Stony Brook University
}

\date[2018]{May 1, 2018}

\begin{document}

\maketitle

\begin{frame}[t]{Problem overview}
  \begin{itemize}
    \item \texttt{TensorFlow} is an open source framework for building neural
      networks.
    \item It provides an interface for creating new \texttt{Optimizer}
      objects which perform different minimization algorithms.
    \item Typical existing implementations are first order descent
      algorithms, but it lacks\footnotemark any implementation of a BFGS-type
      algorithm.
  \end{itemize}
  \footnotetext[1]{See: \url{https://github.com/tensorflow/tensorflow/issues/446}}
\end{frame}

\begin{frame}[t]{Our contribution}
  \begin{itemize}
    \item We implemented stochastic L-BFGS in \texttt{TensorFlow} following the
      work of Wang, Ma, Goldfarb, and Liu (2017).
    \item The approach is based on the iteration step:
      \[
        x_{k+1} = x_k - \alpha_k H_k g_k,
      \]
    \item $\alpha_k$ is the step size,
    \item $H_k$ is the L-BFGS approximate Hessian,
    \item $g_k$ is a batch gradient defined as
      \begin{equation*}
        g_k = \frac{1}{m_k} \sum_{i=1}^{m_k} g(x_k, \xi_{k,i}),
      \end{equation*}
    \item $m_k$ is the batch size,
    \item $\xi_{k,i}$ is randomly sampled training data.
  \end{itemize}
\end{frame}

\begin{frame}[t]{Our contribution}
  \begin{itemize}
    \item The specific approach to approximating the Hessian differs from
      traditional L-BFGS implementations.
    \item Wang, et al showed that convergence (in expectation) is guaranteed by
      applying dampening at specific places in the approximation.
    \item Crucially, this holds even if the problem is \textbf{non-convex}.
    \item Their update, \texttt{SdLBFGS}, 
  \end{itemize}<++>

\end{frame}<++>

\begin{frame}[t]{Implementation}
  \begin{itemize}
    \item Implementing L-BFGS proved challenging, and we discovered why it wasn't
      part of the library already!
    \item Following the advice from the discussion of 
      \href{https://github.com/tensorflow/tensorflow/issues/446}{Issue 446}, we
      implemented L-BFGS as an \texttt{ExternalOptimizerInterface}.
    \item This allowed us to write \texttt{NumPy} code but ``plug in'' to
      \texttt{TensorFlow}.
  \end{itemize}
\end{frame}

\end{document}

### 2018-04-30 Functions.py

@@ -1,38 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Thu Apr 19 21:38:41 2018
-
-@author: vbigmouse
-"""
-
-import numpy as np
-import random as rand
-
-class Functions():
-    def __init__(self, func, datas):
-        self.func = func
-        self.datas = datas
-
-
-    def get_st_value(self, weight, sample_ind):
-
-        avg_value = None
-
-        for ind in sample_ind:
-            fvalue = self.func(weight, self.datas[ind:,-1], 0)
-            avg_value += fvalue / len(sample_ind)
-
-        return avg_value
-
-    def get_st_subgradient(self, weight, sample_ind):
-
-        avg_gradient = None
-        avg_value = None
-
-        for ind in sample_ind:
-            fvalue, gradient = self.func(weight, self.datas[ind:,-1], 1)
-            avg_gradient += gradient / len(sample_ind)
-            avg_value += fvalue / len(sample_ind)
-
-        return fvalue, avg_gradient
-

### 2018-04-30 SDLBFGS.py

@@ -1,157 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Thu Apr 19 02:07:39 2018
-
-@author: vbigmouse
-"""
-import numpy as np
-import random as rand
-import Functions as Fun
-from scipy.optimize import OptimizeResult
-
-class SDLBFGS():
-    def __init__(self, func, datas, *args, 
-            max_iteration=1000, 
-            mem_size=10, 
-            batch_size=50, 
-            step_size=0.1,
-            delta=0.01, 
-            **kwargs):
-        self.max_iteration = max_iteration
-        self.mem_size = mem_size
-        self.batch_size = batch_size
-        self.iteration = 0
-        self.step_size = step_size
-        self.delta = delta
-
-
-        self.Func = Fun(func, datas)
-        self.num_sample = datas.shape[0]
-        self.num_feature = datas.shape[1] - 1
-        self.avg_func_value = None
-
-        self.st_grad = np.zeros(num_feature,1)
-        self.old_st_grad = np.zeros(num_feature,1)
-        self.weight = np.random.rand(num_feature,1)
-        self.old_weight = np.zeros(num_feature,1)
-
-
-        '''S is the update of weight'''
-        self.S = np.zeros((self.num_feature, mem_size))
-        '''Y is diff bewteen st_gradient'''
-        self.Y = np.zeros((self.num_feature, mem_size))
-
-        self.result = OptimizeResult()
-
-    def reset_S_Y(self):
-        self.S = np.zeros((self.num_feature, mem_size))
-        self.Y = np.zeros((self.num_feature, mem_size))
-
-    def save_S_Y(self, s, y):
-#        ind = self.iteration % self.mem_size
-        if self.iteration < self.mem_size:
-            self.S[:, self.iteration] = s
-            self.Y[:, self.iteration] = y
-        else:
-            self.S = np.concatenate(self.S, s, axis=1)
-            self.Y = np.concatenate(self.Y, y, axis=1)
-            self.S = numpy.delete(self.S,(0), axis=1)
-            self.Y = numpy.delete(self.Y,(0), axis=1)
-
-
-
-
-    def update_weight(self):
-
-        ''' calculate stochastic gradient g_{k-1} '''
-        sample_ind = rand.sample(range(1, self.Func.num_sample), self.batch_size)
[…]

### 2018-04-30 algorithms.py

@@ -1,261 +1,0 @@
-######
-######  This file includes different optimization methods
-######
-
-import time
-import numpy as np
-
-
-def bisection( one_d_fun, MIN, MAX, eps=1e-5, maximum_iterations=65536 ):
-    """
-    Identify a critical point of a single-variable real-valued function on a
-    specified interval. 
-    one_d_fun:          a tuple of the form (f, f'), where f : R -> R is differentiable 
-    MIN:                lower bound on the interval on which searching occurs 
-    MAX:                upper bound on the interval on which searching occurs 
-    eps:                the maximum allowed error in the resulting stepsize t
-    maximum_iterations: the maximum allowed number of iterations
-    """
-
-    # counting the number of iterations
-    iterations = 0
-
-    if eps <= 0:
-        raise ValueError("Epsilon must be positive")
-
-    while True:
-
-        iterations += 1
-
-        MID = ( MAX + MIN ) / 2
-
-        # Oracle access to the function value and derivative
-        value, derivative = one_d_fun( MID, 1 )
-
-        if np.abs(derivative) < eps or (MAX - MIN) < 2 * eps:
-            break
-        elif derivative < 0.:
-            MIN = MID
-        else:
-            MAX = MID
-
-        iterations += 1
-        if iterations>=maximum_iterations:
-            break
-
-    return MID
-
-
-def function_on_line(func, x, direction, order):
-    if order==0:
-        value = func(x, order)
-        return value
-    elif order==1:
-        value, gradient = func(x, order)
-        return (value, gradient.T*direction)
-    elif order==2:
-        value, gradient, hessian = func(x, order)
-        return (value, gradient.T*direction, direction.T*hessian*direction)
-    else:
-        raise ValueError("The argument \"order\" should be 0, 1 or 2")
-
-
-###############################################################################
-def exact_line_search( func, x, direction, eps=1e-9, maximum_iterations=65536 ):
-    """ 
-    'Exact' linesearch (using bisection method)
-    func:               the function to optimize It is called as "value, gradient = func( x, 1 )
-    x:                  the current iterate
-    direction:          the direction along which to perform the linesearch
-    eps:                the maximum allowed error in the resulting stepsize t
-    maximum_iterations: the maximum allowed number of iterations
-    """
-
-    x = np.matrix( x )
-    direction = np.matrix( direction )
-
-    value_0 = func( x , 0 )
-
[…]

### 2018-04-30 cocob/mnist/mnist_fully_connected.py

@@ -1,111 +1,0 @@
-# Copyright 2015 The TensorFlow Authors. All Rights Reserved.
-#
-# Licensed under the Apache License, Version 2.0 (the "License");
-# you may not use this file except in compliance with the License.
-# You may obtain a copy of the License at
-#
-#     http://www.apache.org/licenses/LICENSE-2.0
-#
-# Unless required by applicable law or agreed to in writing, software
-# distributed under the License is distributed on an "AS IS" BASIS,
-# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-# See the License for the specific language governing permissions and
-# limitations under the License.
-# ==============================================================================
-
-# Modifications copyright (C) 2017 Francesco Orabona, Tatiana Tommasi 
-
-"""A fully connected MNIST classifier using convolutional layers.
-
-See extensive documentation at
-https://www.tensorflow.org/get_started/mnist/pros
-"""
-# Disable linter warnings to maintain consistency with tutorial.
-# pylint: disable=invalid-name
-# pylint: disable=g-bad-import-order
-
-from __future__ import absolute_import
-from __future__ import division
-from __future__ import print_function
-
-import argparse
-import sys
-
-import input_data
-
-import tensorflow as tf
-import numpy as np
-
-sys.path.insert(0, '../optimizer/')
-
-import sqn_optimizer
-
-FLAGS = None
-
-
-def weight_variable(shape):
-  """weight_variable generates a weight variable of a given shape."""
-  initial = tf.truncated_normal(shape, stddev=0.1)
-  return tf.Variable(initial)
-
-
-def bias_variable(shape):
-  """bias_variable generates a bias variable of a given shape."""
-  initial = tf.constant(0.1, shape=shape)
-  return tf.Variable(initial)
-
-
-def main(_):
-  # Import data
-  mnist = input_data.read_data_sets('data', one_hot=True, validation_size=0)
-
-  # Create the model
-  x = tf.placeholder(tf.float32, [None, 784])
-
-  # Define loss and optimizer
-  y_ = tf.placeholder(tf.float32, [None, 10])
-
-  W_fc1 = weight_variable([28*28, 1000])
-  b_fc1 = bias_variable([1000])
-  h_fc1 = tf.nn.relu(tf.matmul(x, W_fc1) + b_fc1)
[…]

### 2018-04-30 cocob/optimizer/__init__.py

@@ -1,1 +1,0 @@
-from sqn_optimizer import *

### 2018-04-30 cocob/optimizer/sqn_optimizer.py

@@ -1,99 +1,0 @@
-# Copyright 2017 Francesco Orabona. All Rights Reserved.
-#
-# Licensed under the Apache License, Version 2.0 (the "License");
-# you may not use this file except in compliance with the License.
-# You may obtain a copy of the License at
-#
-#     http://www.apache.org/licenses/LICENSE-2.0
-#
-# Unless required by applicable law or agreed to in writing, software
-# distributed under the License is distributed on an "AS IS" BASIS,
-# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-# See the License for the specific language governing permissions and
-# limitations under the License.
-# ==============================================================================
-
-
-'''
-Stochastic Quasi Newton (SQN) optimizer
-'''
-
-from tensorflow.python.framework import ops
-from tensorflow.python.ops import state_ops
-from tensorflow.python.ops import control_flow_ops
-from tensorflow.python.framework import constant_op
-from tensorflow.python.training.optimizer import Optimizer
-import tensorflow as tf
-
-
-class SQN(Optimizer):
-    def __init__(self, alpha=100, mem_size=10, batch_size=1, step_size=1., 
-            use_locking=False, name='SQN'):
-        '''
-        constructs a new SQN optimizer
-        '''
-        super(SQN, self).__init__(use_locking, name)
-        self._alpha = alpha
-        self._mem_size = mem_size
-        self._batch_size = batch_size
-        self._step_size = step_size
-
-    def _create_slots(self, var_list):
-        # var_list is any collection of features or other useful data
-        for v in var_list:
-            with ops.colocate_with(v):
-                gradients_sum = constant_op.constant(0, 
-                                                     shape=v.get_shape(),
-                                                     dtype=v.dtype.base_dtype)
-                grad_norm_sum = constant_op.constant(0, 
-                                                     shape=v.get_shape(),
-                                                     dtype=v.dtype.base_dtype)
-                L = constant_op.constant(1e-8, shape=v.get_shape(), dtype=v.dtype.base_dtype)
-                tilde_w = constant_op.constant(0.0, shape=v.get_shape(), dtype=v.dtype.base_dtype)
-                reward = constant_op.constant(0.0, shape=v.get_shape(), dtype=v.dtype.base_dtype)
-
[…]

### 2018-04-30 cocob/optimizer/sqn_optimizer_hooks.py

@@ -1,185 +1,0 @@
-
-# Copyright 2016 The TensorFlow Authors. All Rights Reserved.
-#
-# Licensed under the Apache License, Version 2.0 (the "License");
-# you may not use this file except in compliance with the License.
-# You may obtain a copy of the License at
-#
-#     http://www.apache.org/licenses/LICENSE-2.0
-#
-# Unless required by applicable law or agreed to in writing, software
-# distributed under the License is distributed on an "AS IS" BASIS,
-# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-# See the License for the specific language governing permissions and
-# limitations under the License.
-# ==============================================================================
-"""TensorFlow interface for third-party optimizers."""
- 
-from __future__ import absolute_import
-from __future__ import division
-from __future__ import print_function
- 
-import numpy as np
- 
-from tensorflow.python.framework import ops
-from tensorflow.python.ops import array_ops
-from tensorflow.python.ops import gradients
-from tensorflow.python.ops import variables
-from tensorflow.python.platform import tf_logging as logging
-from tensorflow.python.framework import constant_op
-from tensorflow.python.training.optimizer import ExternalOptimizerInterface
-import tensorflow as tf
-
-
-class SQNOptimizer(ExternalOptimizerInterface):
-  """Wrapper allowing `scipy.optimize.minimize` to operate a `tf.Session`.
- 
-  Example:
- 
-  ```python
-  vector = tf.Variable([7., 7.], 'vector')
- 
-  # Make vector norm as small as possible.
-  loss = tf.reduce_sum(tf.square(vector))
- 
-  optimizer = ScipyOptimizerInterface(loss, options={'maxiter': 100})
- 
-  with tf.Session() as session:
-    optimizer.minimize(session)
- 
-  # The value of vector should now be [0., 0.].
-  ```
- 
-  Example with simple bound constraints:
- 
-  ```python
-  vector = tf.Variable([7., 7.], 'vector')
- 
-  # Make vector norm as small as possible.
-  loss = tf.reduce_sum(tf.square(vector))
- 
-  optimizer = ScipyOptimizerInterface(
-      loss, var_to_bounds={vector: ([1, 2], np.infty)})
- 
-  with tf.Session() as session:
-    optimizer.minimize(session)
- 
-  # The value of vector should now be [1., 2.].
-  ```
- 
[…]

### 2018-04-30 hw1_functions.py

@@ -1,125 +1,0 @@
-######
-######  This file includes different functions used in HW1
-######
-
-import numpy as np
-import matplotlib.pyplot as plt
-import time
-from math import exp
-
-def weird_func( x, order=0 ):
-
-  # f(x) = x^4 + 6x^2 + 12(x-4)e^(x-1)
-  value = pow(x, 4) + 6 * pow(x, 2) + 12 * (x - 4) * exp(x - 1)
-  
-  if order==0:
-      return value
-  elif order==1:
-      # f'(x) = 4x^3 + 12x + 12(x-3)e^(x-1)
-      gradient = 4 * pow(x, 3) + 12 * x + 12 * (x - 3) * exp(x - 1)
-
-      return (value, gradient)
-  elif order==2:
-      # f'(x) = 4x^3 + 12x + 12(x-3)e^(x-1)
-      gradient = 4 * pow(x, 3) + 12 * x + 12 * (x - 3) * exp(x - 1)
-
-      # f''(x)= 12 (1 + e^(-1 + x) (-2 + x) + x^2)
-      hessian = 12 * (1 + (x-2) * exp(x-1) + pow(x,2))
-
-      return (value, gradient, hessian)
-  else:
-        raise ValueError("The argument \"order\" should be 0, 1 or 2")
-
-
-###############################################################################
-def boyd_example_func(x, order=0):
-  a=np.matrix('1  3')
-  b=np.matrix('1  -3')
-  c=np.matrix('-1  0')
-  x=np.asmatrix(x)
-  
-  value = exp(a*x-0.1)+exp(b*x-0.1)+exp(c*x-0.1)
-  if order==0:
-      return value
-  elif order==1:
-      gradient = a.T*exp(a*x-0.1)+b.T*exp(b*x-0.1)+c.T*exp(c*x-0.1)
-      return (value, gradient)
-  elif order==2:
-      gradient = a.T*exp(a*x-0.1)+b.T*exp(b*x-0.1)+c.T*exp(c*x-0.1)
-      hessian = a.T*a*exp(a*x-0.1)+b.T*b*exp(b*x-0.1)+c.T*c*exp(c*x-0.1)
-      return (value, gradient, hessian)
-  else:
-        raise ValueError("The argument \"order\" should be 0, 1 or 2")
-
-
-###############################################################################
-def quadratic( H, b, x, order=0 ):
-    """ 
-    Quadratic Objective
-    H:          the Hessian matrix
-    b:          the vector of linear coefficients
-    x:          the current iterate
[…]

### 2018-04-30 sqn/mnist/external_optimizer.py

@@ -1,0 +1,335 @@
+# Copyright 2016 Google Inc. All Rights Reserved.
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+# ==============================================================================
+
+"""TensorFlow interface for third-party optimizers."""
+
+from __future__ import absolute_import
+from __future__ import division
+from __future__ import print_function
+
+from tensorflow.python.framework import ops
+from tensorflow.python.ops import array_ops
+from tensorflow.python.ops import gradients
+from tensorflow.python.ops import variables
+
+
+__all__ = ['ExternalOptimizerInterface', 'ScipyOptimizerInterface']
+
+
+class ExternalOptimizerInterface(object):
+  """Base class for interfaces with external optimization algorithms.
+
+  Subclass this and implement `_minimize` in order to wrap a new optimization
+  algorithm.
+
+  `ExternalOptimizerInterface` should not be instantiated directly; instead use
+  e.g. `ScipyOptimizerInterface`.
+
+  @@__init__
+
+  @@minimize
+  """
+
+  def __init__(self, loss, var_list=None, equalities=None, inequalities=None,
+               **optimizer_kwargs):
+    """Initialize a new interface instance.
+
+    Args:
+      loss: A scalar `Tensor` to be minimized.
+      var_list: Optional list of `Variable` objects to update to minimize
+        `loss`.  Defaults to the list of variables collected in the graph
+        under the key `GraphKeys.TRAINABLE_VARIABLES`.
+      equalities: Optional list of equality constraint scalar `Tensor`s to be
+        held equal to zero.
+      inequalities: Optional list of inequality constraint scalar `Tensor`s
+        to be kept nonnegative.
+      **optimizer_kwargs: Other subclass-specific keyword arguments.
+    """
+    self._loss = loss
+    self._equalities = equalities or []
+    self._inequalities = inequalities or []
+
+    if var_list is None:
[…]

### 2018-04-30 sqn/mnist/mnist_fully_connected.py

@@ -1,0 +1,119 @@
+# Copyright 2015 The TensorFlow Authors. All Rights Reserved.
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+# ==============================================================================
+
+# Modifications copyright (C) 2017 Francesco Orabona, Tatiana Tommasi 
+
+"""A fully connected MNIST classifier using convolutional layers.
+
+See extensive documentation at
+https://www.tensorflow.org/get_started/mnist/pros
+"""
+# Disable linter warnings to maintain consistency with tutorial.
+# pylint: disable=invalid-name
+# pylint: disable=g-bad-import-order
+
+from __future__ import absolute_import
+from __future__ import division
+from __future__ import print_function
+
+import argparse
+import sys
+
+import input_data
+
+import tensorflow as tf
+import numpy as np
+
+sys.path.insert(0, '../optimizer/')
+
+
+from external_optimizer import ScipyOptimizerInterface
+Optimizer = ScipyOptimizerInterface
+
+FLAGS = None
+
+
+def weight_variable(shape):
+    """weight_variable generates a weight variable of a given shape."""
+    initial = tf.truncated_normal(shape, stddev=0.1)
+    return tf.Variable(initial)
+
+
+def bias_variable(shape):
+    """bias_variable generates a bias variable of a given shape."""
+    initial = tf.constant(0.1, shape=shape)
+    return tf.Variable(initial)
+
+
+def main(_):
+    # Import data
+    mnist = input_data.read_data_sets('data', one_hot=True, validation_size=0)
+
+    # Create the model
+    x = tf.placeholder(tf.float32, [None, 784])
+
+    # Define loss and optimizer
+    y_ = tf.placeholder(tf.float32, [None, 10])
+
+    W_fc1 = weight_variable([28*28, 1000])
+    b_fc1 = bias_variable([1000])
+    h_fc1 = tf.nn.relu(tf.matmul(x, W_fc1) + b_fc1)
+
[…]

### 2018-04-30 sqn/mnist/mnist_softmax.py

@@ -1,0 +1,82 @@
+# Copyright 2015 The TensorFlow Authors. All Rights Reserved.
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+# ==============================================================================
+
+"""A very simple MNIST classifier.
+
+See extensive documentation at
+http://tensorflow.org/tutorials/mnist/beginners/index.md
+"""
+from __future__ import absolute_import
+from __future__ import division
+from __future__ import print_function
+
+import argparse
+import sys
+
+from tensorflow.examples.tutorials.mnist import input_data
+from tensorflow.contrib.opt.python.training.external_optimizer import ScipyOptimizerInterface
+from tensorflow.python.ops import variables
+
+import tensorflow as tf
+
+FLAGS = None
+
+
+def main(_):
+  # Import data
+  mnist = input_data.read_data_sets(FLAGS.data_dir, one_hot=True)
+
+  # Create the model
+  x = tf.placeholder(tf.float32, [None, 784])
+  W = tf.Variable(tf.zeros([784, 10]))
+  b = tf.Variable(tf.zeros([10]))
+  y = tf.matmul(x, W) + b
+
+  # Define loss and optimizer
+  y_ = tf.placeholder(tf.float32, [None, 10])
+
+  # The raw formulation of cross-entropy,
+  #
+  #   tf.reduce_mean(-tf.reduce_sum(y_ * tf.log(tf.nn.softmax(y)),
+  #                                 reduction_indices=[1]))
+  #
+  # can be numerically unstable.
+  #
+  # So here we use tf.nn.softmax_cross_entropy_with_logits on the raw
+  # outputs of 'y', and then average across the batch.
+  cross_entropy = tf.reduce_mean(
+      tf.nn.softmax_cross_entropy_with_logits(labels=y_, logits=y))
+  #train_step = tf.train.GradientDescentOptimizer(0.5).minimize(cross_entropy)
[…]

### 2018-04-30 sqn/optimizer/__init__.py

@@ -1,0 +1,1 @@
+from sqn_optimizer_hooks import SQNOptimizer

### 2018-04-30 sqn/optimizer/sd_lbfgs.py

@@ -1,0 +1,326 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Apr 19 02:07:39 2018
+
+@author: vbigmouse
+"""
+from numpy import dot, zeros, asarray
+from scipy.optimize import OptimizeResult
+
+
+def harmonic_sequence(k):
+    # For any nonnegative integer
+    return 1./(k+1)
+
+class SdLBFGS():
+    def __init__(self, func, initial_val, *args,
+            max_iterations=1000,
+            mem_size=10,
+            batch_size=50,
+            step_size=harmonic_sequence,
+            delta=0.01,
+            tol=1e-5,
+            **kwargs):
+        
+        self._func = func
+        self._initial_val = asarray(initial_val)
+        self._max_iterations = max_iterations
+        self._max_mem_size = mem_size
+        self._mem_size = 0
+        self._batch_size = batch_size
+        self._step_size = step_size
+        self._delta = delta
+        self._tolerance = tol
+
+        self._iterations = 0
+        self._current_val = self._initial_val.copy()
+        self._previous_val = zeros(self._initial_val.shape)
+        self._previous_grad = zeros(self._initial_val.shape)
+
+        self._backward_errors = ShiftList(self._max_mem_size)
+        self._ybars = ShiftList(self._max_mem_size)
+        self._rhos = ShiftList(self._max_mem_size)
+
+    
+    def run(self):
+        while self._iterations < self._max_iterations:
+            self.sqn_step()
+        return self.result()
+
+
+    def sqn_step(self):
+        # compute stochastic gradient
+        grad = zeros(self._current_val.shape)
+        for i in range(self._batch_size):
+            _, g = self._func(self._current_val)
+            grad += g
+            print(g)
+        grad /= self._batch_size
+
+        # stopping criterion, sends back to self.run()
+        if dot(grad, grad) <= self._tolerance:
+            return
+
+        # compute search direction
+        direction = self.sdlbfgs_step(grad)
+
+        # update the iteration invariants
+        self._previous_val = self._current_val.copy()
+        self._current_val -= self._step_size(self._iterations) * direction
+        self._previous_grad = grad.copy()
+        self._iterations += 1
+
+
+    def sdlbfgs_step(self, stochastic_grad):
+        p = self._mem_size
[…]

### 2018-04-30 sqn/optimizer/sqn_optimizer_hooks.py

@@ -1,0 +1,174 @@
+
+# Copyright 2016 The TensorFlow Authors. All Rights Reserved.
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+# ==============================================================================
+"""TensorFlow interface for third-party optimizers."""
+ 
+from __future__ import absolute_import
+from __future__ import division
+from __future__ import print_function
+ 
+import numpy as np
+ 
+from tensorflow.python.framework import ops
+from tensorflow.python.ops import array_ops
+from tensorflow.python.ops import gradients
+from tensorflow.python.ops import variables
+from tensorflow.python.platform import tf_logging as logging
+from tensorflow.python.framework import constant_op
+from tensorflow.contrib.opt.python.training.external_optimizer import ExternalOptimizerInterface
+import tensorflow as tf
+
+from sd_lbfgs import SdLBFGS
+
+
+class SQNOptimizer(ExternalOptimizerInterface):
+  """Wrapper allowing `scipy.optimize.minimize` to operate a `tf.Session`.
+ 
+  Example:
+ 
+  ```python
+  vector = tf.Variable([7., 7.], 'vector')
+ 
+  # Make vector norm as small as possible.
+  loss = tf.reduce_sum(tf.square(vector))
+ 
+  optimizer = ScipyOptimizerInterface(loss, options={'maxiter': 100})
+ 
+  with tf.Session() as session:
+    optimizer.minimize(session)
+ 
+  # The value of vector should now be [0., 0.].
+  ```
+ 
+  Example with simple bound constraints:
+ 
+  ```python
+  vector = tf.Variable([7., 7.], 'vector')
+ 
+  # Make vector norm as small as possible.
+  loss = tf.reduce_sum(tf.square(vector))
+ 
+  optimizer = ScipyOptimizerInterface(
+      loss, var_to_bounds={vector: ([1, 2], np.infty)})
+ 
+  with tf.Session() as session:
+    optimizer.minimize(session)
+ 
[…]

### 2018-04-30 sqn/optimizer/test.py

@@ -1,0 +1,30 @@
+#import tensorflow as tf
+from sqn_optimizer_hooks import SQNOptimizer
+from sd_lbfgs import SdLBFGS
+from hw1_functions import weird_func
+import numpy as np
+from time import sleep
+from pprint import pprint
+
+optimizer = SdLBFGS(lambda x: weird_func(x, order=1), 0.)
+
+for i in range(5):
+    pprint(optimizer.__dict__)
+    optimizer.sqn_step()
+
+# vector = tf.Variable([7., 7.], 'vector')
+# 
+# # Make vector norm as small as possible.
+# loss = tf.reduce_sum(tf.square(vector))
+# 
+# print(loss)
+# 
+# optimizer = SQNOptimizer(loss)
+#  
+# with tf.Session() as session:
+#     print("Inside the session.")
+#     session.run(tf.global_variables_initializer())
+# 
+#     optimizer.minimize(session)
+# 
+#     # The value of vector should now be [0., 0.].

### 2018-04-30 sqn/optimizer/hw1_functions.py
More updates

@@ -30,7 +30,6 @@ def weird_func( x, order=0 ):
   else:
         raise ValueError("The argument \"order\" should be 0, 1 or 2")
 
-
 ###############################################################################
 def boyd_example_func(x, order=0):
   a=np.matrix('1  3')

### 2018-04-30 sqn/optimizer/sd_lbfgs.py

@@ -10,7 +10,7 @@ from scipy.optimize import OptimizeResult
 
 def harmonic_sequence(k):
     # For any nonnegative integer
-    return 1./(k+1)
+    return 1./(10 * (k+1))
 
 class SdLBFGS():
     def __init__(self, func, initial_val, *args,
@@ -35,6 +35,7 @@ class SdLBFGS():
         self._iterations = 0
         self._current_val = self._initial_val.copy()
         self._previous_val = zeros(self._initial_val.shape)
+        self._current_grad = zeros(self._initial_val.shape)
         self._previous_grad = zeros(self._initial_val.shape)
 
         self._backward_errors = ShiftList(self._max_mem_size)
@@ -54,31 +55,31 @@ class SdLBFGS():
         for i in range(self._batch_size):
             _, g = self._func(self._current_val)
             grad += g
-            print(g)
         grad /= self._batch_size
+        self._current_grad = grad
 
         # stopping criterion, sends back to self.run()
-        if dot(grad, grad) <= self._tolerance:
+        if dot(self._current_grad, self._current_grad) <= self._tolerance:
             return
 
         # compute search direction
-        direction = self.sdlbfgs_step(grad)
+        direction = self.sdlbfgs_step()
 
         # update the iteration invariants
         self._previous_val = self._current_val.copy()
         self._current_val -= self._step_size(self._iterations) * direction
-        self._previous_grad = grad.copy()
+        self._previous_grad = self._current_grad.copy()
         self._iterations += 1
 
 
-    def sdlbfgs_step(self, stochastic_grad):
-        p = self._mem_size
+    def sdlbfgs_step(self):
+        #print(f'{self._current_val}, {self._current_grad}')
         k = self._iterations
 
         s = self._current_val - self._previous_val
-        y = stochastic_grad - self._previous_grad
+        y = self._current_grad - self._previous_grad
 
-        print(f's: {s}, y: {y}')
+        #print(f's: {s}, y: {y}')
 
         # compute theta
         sTy = dot(s, y)
@@ -97,23 +98,28 @@ class SdLBFGS():
         self.update_memory_vars(s, y_bar, rho)
 
         if k == 0:
-            return stochastic_grad
+            return self._current_grad
 
-        u = stochastic_grad.copy()
+        u = self._current_grad.copy()
         mus = []
+
+        p = self._mem_size
+        print(p)
[…]

### 2018-04-30 sqn/optimizer/test.py

@@ -6,10 +6,18 @@ import numpy as np
 from time import sleep
 from pprint import pprint
 
-optimizer = SdLBFGS(lambda x: weird_func(x, order=1), 0.)
+def quadratic(x, order=1):
+    if order == 0:
+        return x * x
+    if order==1:
+        return x * x, 2 * x
 
-for i in range(5):
-    pprint(optimizer.__dict__)
+
+optimizer = SdLBFGS(lambda x: quadratic(x, order=1), -1.,
+        batch_size=2)
+
+for i in range(50):
+    # pprint(optimizer.__dict__)
     optimizer.sqn_step()
 
 # vector = tf.Variable([7., 7.], 'vector')

### 2018-04-30 sqn/optimizer/sd_lbfgs.py
Merge

@@ -21,7 +21,7 @@ class SdLBFGS():
             delta=0.01,
             tol=1e-5,
             **kwargs):
-        
+
         self._func = func
         self._initial_val = asarray(initial_val)
         self._max_iterations = max_iterations
@@ -41,7 +41,7 @@ class SdLBFGS():
         self._ybars = ShiftList(self._max_mem_size)
         self._rhos = ShiftList(self._max_mem_size)
 
-    
+
     def run(self):
         while self._iterations < self._max_iterations:
             self.sqn_step()
@@ -54,7 +54,7 @@ class SdLBFGS():
         for i in range(self._batch_size):
             _, g = self._func(self._current_val)
             grad += g
-            print(g)
+#            print(g)
         grad /= self._batch_size
 
         # stopping criterion, sends back to self.run()
@@ -84,7 +84,7 @@ class SdLBFGS():
         sTy = dot(s, y)
         yTy = dot(y, y)
         sTs = dot(s, s)
-        
+
         gamma = max(yTy / sTy, self._delta)
 
         if sTy < 0.25 * gamma * sTs:
@@ -141,9 +141,9 @@ class SdLBFGS():
         self.result['jac'] = g
         self.result['nfev'] = self._batch_size * self._iterations
         self.result['njev'] = self._batch_size * self._iterations
-        self.result['nit'] = self._iterations 
+        self.result['nit'] = self._iterations
 
-        return result       
+        return result
 
 
 
@@ -176,12 +176,12 @@ class ShiftList():
 
 
 # class SDLBFGS():
-#     def __init__(self, func, datas, *args, 
-#             max_iteration=1000, 
-#             mem_size=10, 
-#             batch_size=50, 
+#     def __init__(self, func, datas, *args,
+#             max_iteration=1000,
+#             mem_size=10,
+#             batch_size=50,
 #             step_size=0.1,
-#             delta=0.01, 
+#             delta=0.01,
 #             **kwargs):
 #         self.max_iteration = max_iteration
 #         self.mem_size = mem_size
@@ -189,30 +189,30 @@ class ShiftList():
 #         self.iteration = 0
 #         self.step_size = step_size
 #         self.delta = delta
-# 
-# 
+#
+#
 #         self.func = func
[…]

### 2018-04-30 sqn/optimizer/test.py

@@ -13,18 +13,18 @@ for i in range(5):
     optimizer.sqn_step()
 
 # vector = tf.Variable([7., 7.], 'vector')
-# 
+#
 # # Make vector norm as small as possible.
 # loss = tf.reduce_sum(tf.square(vector))
-# 
+#
 # print(loss)
-# 
+#
 # optimizer = SQNOptimizer(loss)
-#  
+#
 # with tf.Session() as session:
 #     print("Inside the session.")
 #     session.run(tf.global_variables_initializer())
-# 
+#
 #     optimizer.minimize(session)
-# 
+#
 #     # The value of vector should now be [0., 0.].

### 2018-04-30 sqn/optimizer/sd_lbfgs.py
update index of mu

@@ -4,13 +4,13 @@ Created on Thu Apr 19 02:07:39 2018
 
 @author: vbigmouse
 """
-from numpy import dot, zeros, asarray
+from numpy import dot, zeros, asarray, sqrt
 from scipy.optimize import OptimizeResult
 
 
 def harmonic_sequence(k):
     # For any nonnegative integer
-    return 1./(10 * (k+1))
+    return 0.1/(sqrt(k+1))
 
 class SdLBFGS():
     def __init__(self, func, initial_val, *args,
@@ -52,11 +52,13 @@ class SdLBFGS():
     def sqn_step(self):
         # compute stochastic gradient
         grad = zeros(self._current_val.shape)
+#        print(self._current_val)
         for i in range(self._batch_size):
             _, g = self._func(self._current_val)
             grad += g
         grad /= self._batch_size
         self._current_grad = grad
+
 
         # stopping criterion, sends back to self.run()
         if dot(self._current_grad, self._current_grad) <= self._tolerance:
@@ -64,6 +66,7 @@ class SdLBFGS():
 
         # compute search direction
         direction = self.sdlbfgs_step()
+        print(f'  direction {direction}')
 
         # update the iteration invariants
         self._previous_val = self._current_val.copy()
@@ -78,8 +81,8 @@ class SdLBFGS():
 
         s = self._current_val - self._previous_val
         y = self._current_grad - self._previous_grad
-
-        #print(f's: {s}, y: {y}')
+        print(f'current val: {self._current_val}')
+        print(f's: {s}, y: {y}')
 
         # compute theta
         sTy = dot(s, y)
@@ -92,32 +95,37 @@ class SdLBFGS():
             theta = 0.75 * gamma * sTs / (gamma * sTs - sTy)
         else:
             theta = 1.
-
+        #theta =1
         y_bar = theta * y + ((1. - theta) * gamma) * s
         rho = 1. / dot(s, y_bar)
-        self.update_memory_vars(s, y_bar, rho)
+
 
         if k == 0:
             return self._current_grad
+        self.update_memory_vars(s, y_bar, rho)
 
         u = self._current_grad.copy()
         mus = []
 
         p = self._mem_size
-        print(p)
+
+        print(self._rhos)
         irange = range(min(p, k-1))
         for i in irange:
             mu = self._rhos[k-i-1] * dot(u, self._backward_errors[k-i-1])
[…]

### 2018-04-30 sqn/optimizer/test.py

@@ -11,14 +11,16 @@ def quadratic(x, order=1):
         return x * x
     if order==1:
         return x * x, 2 * x
+#%%
 
-
-optimizer = SdLBFGS(lambda x: quadratic(x, order=1), -1.,
+optimizer = SdLBFGS(lambda x: quadratic(x, order=1), 1.,
         batch_size=2)
 
-for i in range(50):
-    # pprint(optimizer.__dict__)
+for i in range(10):
+    print(i)
+#    pprint(optimizer.__dict__)
     optimizer.sqn_step()
+    print()
 
 # vector = tf.Variable([7., 7.], 'vector')
 #

### 2018-04-30 sqn/optimizer/sd_lbfgs.py
Auto stash before merge of "master" and "origin/master"

@@ -78,6 +78,7 @@ class SdLBFGS():
     def sdlbfgs_step(self):
         #print(f'{self._current_val}, {self._current_grad}')
         k = self._iterations
+        p = self._mem_size
 
         s = self._current_val - self._previous_val
         y = self._current_grad - self._previous_grad
@@ -107,27 +108,19 @@ class SdLBFGS():
         u = self._current_grad.copy()
         mus = []
 
-        p = self._mem_size
-
-        print(self._rhos)
         irange = range(min(p, k-1))
+        print(irange, len(mus))
         for i in irange:
             mu = self._rhos[k-i-1] * dot(u, self._backward_errors[k-i-1])
             u -= mu * self._ybars[k-i-1]
             mus.append(mu)
-        print(f' !!!mus {mus}')
 
         v = (1. / gamma) * u
         for i in irange:
-            nu = self._rhos[k-p+i] * \
-                    dot(v, self._ybars[k-p+i])
+            print(p, i)
+            nu = self._rhos[k-p+i] * dot(v, self._ybars[k-p+i])
+            v += (mus[p-i-1] - nu) * self._backward_errors[k-p+i]
 
-#            print(f"p-i-1: {p-i-1}")
-#            print(f"size mus: {len(mus)}")
-
-            v += (mus[p-i-2] - nu) * self._backward_errors[k-p+i]
-
-        #print(f'v: {v}')
         return v
 
 

### 2018-04-30 sqn/optimizer/sd_lbfgs.py
The code is working.... I don't know why.

@@ -10,7 +10,7 @@ from scipy.optimize import OptimizeResult
 
 def harmonic_sequence(k):
     # For any nonnegative integer
-    return 0.1/(sqrt(k+1))
+    return 0.9/(sqrt(k+1))
 
 class SdLBFGS():
     def __init__(self, func, initial_val, *args,
@@ -19,7 +19,7 @@ class SdLBFGS():
             batch_size=50,
             step_size=harmonic_sequence,
             delta=0.01,
-            tol=1e-5,
+            tol=1e-4,
             **kwargs):
 
         self._func = func
@@ -42,9 +42,15 @@ class SdLBFGS():
         self._ybars = ShiftList(self._max_mem_size)
         self._rhos = ShiftList(self._max_mem_size)
 
+        self._iteration_vals = [self._current_val]
+
+        self._result = OptimizeResult()
+        self._result['success'] = False
+        
+
 
     def run(self):
-        while self._iterations < self._max_iterations:
+        while not self._result['success'] and self._iterations < self._max_iterations:
             self.sqn_step()
         return self.result()
 
@@ -52,7 +58,6 @@ class SdLBFGS():
     def sqn_step(self):
         # compute stochastic gradient
         grad = zeros(self._current_val.shape)
-#        print(self._current_val)
         for i in range(self._batch_size):
             _, g = self._func(self._current_val)
             grad += g
@@ -61,16 +66,17 @@ class SdLBFGS():
 
 
         # stopping criterion, sends back to self.run()
+        print(dot(self._current_grad, self._current_grad))
         if dot(self._current_grad, self._current_grad) <= self._tolerance:
-            return
+            self._result['success'] = True
 
         # compute search direction
         direction = self.sdlbfgs_step()
-        print(f'  direction {direction}')
 
         # update the iteration invariants
         self._previous_val = self._current_val.copy()
         self._current_val -= self._step_size(self._iterations) * direction
+        self._iteration_vals.append(self._current_val.copy())
         self._previous_grad = self._current_grad.copy()
         self._iterations += 1
 
@@ -82,8 +88,6 @@ class SdLBFGS():
 
         s = self._current_val - self._previous_val
         y = self._current_grad - self._previous_grad
-        print(f'current val: {self._current_val}')
-        print(f's: {s}, y: {y}')
 
         # compute theta
         sTy = dot(s, y)
@@ -109,7 +113,6 @@ class SdLBFGS():
         mus = []
 
[…]

### 2018-04-30 sqn/optimizer/test.py

@@ -13,14 +13,21 @@ def quadratic(x, order=1):
         return x * x, 2 * x
 #%%
 
-optimizer = SdLBFGS(lambda x: quadratic(x, order=1), 1.,
+optimizer = SdLBFGS(lambda x: weird_func(x, order=1), 1.,
         batch_size=2)
 
-for i in range(10):
-    print(i)
-#    pprint(optimizer.__dict__)
-    optimizer.sqn_step()
-    print()
+# for i in range(10):
+#     optimizer.sqn_step()
+
+result = optimizer.run()
+import matplotlib.pyplot as plt
+
+x = np.array(result['iteration_vals'])
+vq = np.vectorize(weird_func)
+y = vq(x, order=0)
+plt.plot(x,y, 'o')
+plt.show()
+
 
 # vector = tf.Variable([7., 7.], 'vector')
 #

### 2018-05-01 sqn/optimizer/sd_lbfgs.py
Fix array dot

@@ -24,6 +24,7 @@ class SdLBFGS():
 
         self._func = func
         self._initial_val = asarray(initial_val)
+#        self._initial_val = self._initial_val.reshape(self._initial_val.shape[0])
         self._max_iterations = max_iterations
         self._max_mem_size = mem_size
         self._mem_size = 0
@@ -46,7 +47,7 @@ class SdLBFGS():
 
         self._result = OptimizeResult()
         self._result['success'] = False
-        
+
 
 
     def run(self):
@@ -66,8 +67,9 @@ class SdLBFGS():
 
 
         # stopping criterion, sends back to self.run()
-        print(dot(self._current_grad, self._current_grad))
-        if dot(self._current_grad, self._current_grad) <= self._tolerance:
+        print(dot(self._current_grad.T, self._current_grad))
+        if dot(self._current_grad.T, self._current_grad) <= self._tolerance:
+
             self._result['success'] = True
 
         # compute search direction
@@ -90,9 +92,9 @@ class SdLBFGS():
         y = self._current_grad - self._previous_grad
 
         # compute theta
-        sTy = dot(s, y)
-        yTy = dot(y, y)
-        sTs = dot(s, s)
+        sTy = dot(s.T, y)
+        yTy = dot(y.T, y)
+        sTs = dot(s.T, s)
 
         gamma = max(yTy / sTy, self._delta)
 
@@ -102,7 +104,7 @@ class SdLBFGS():
             theta = 1.
         #theta =1
         y_bar = theta * y + ((1. - theta) * gamma) * s
-        rho = 1. / dot(s, y_bar)
+        rho = 1. / dot(s.T, y_bar)
 
 
         if k == 0:
@@ -114,13 +116,13 @@ class SdLBFGS():
 
         irange = range(min(p, k-1))
         for i in irange:
-            mu = self._rhos[k-i-1] * dot(u, self._backward_errors[k-i-1])
+            mu = self._rhos[k-i-1] * dot(u.T, self._backward_errors[k-i-1])
             u -= mu * self._ybars[k-i-1]
             mus.append(mu)
 
         v = (1. / gamma) * u
         for i in irange:
-            nu = self._rhos[k-p+i] * dot(v, self._ybars[k-p+i])
+            nu = self._rhos[k-p+i] * dot(v.T, self._ybars[k-p+i])
             v += (mus[p-i-1] - nu) * self._backward_errors[k-p+i]
 
         return v

### 2018-05-01 sqn/optimizer/test.py

@@ -2,6 +2,7 @@
 from sqn_optimizer_hooks import SQNOptimizer
 from sd_lbfgs import SdLBFGS
 from hw1_functions import weird_func
+from hw4_functions import svm_objective_function_stochastic as svm_func
 import numpy as np
 from time import sleep
 from pprint import pprint
@@ -11,20 +12,37 @@ def quadratic(x, order=1):
         return x * x
     if order==1:
         return x * x, 2 * x
+
+#%%
+data = np.loadtxt('HIGGS_subset.csv', delimiter=',')
+
+labels = np.asmatrix(2*data[:,0]-1).T
+features = np.asmatrix(data[:,1:])
+data = np.concatenate((features, labels), axis = 1)
+d = features.shape[1]
+w = np.zeros((d,1))
 #%%
 
-optimizer = SdLBFGS(lambda x: weird_func(x, order=1), 1.,
-        batch_size=2)
+#optimizer = SdLBFGS(lambda x: weird_func(x, order=1), 1., batch_size=2)
+
+
+optimizer = SdLBFGS(lambda x: svm_func(x, order=1, data=data, minibatch_size=50), w,
+        batch_size=50)
 
 # for i in range(10):
 #     optimizer.sqn_step()
 
 result = optimizer.run()
 import matplotlib.pyplot as plt
-
+#%%
+from hw4_functions import svm_objective_function as svm_func_obj
 x = np.array(result['iteration_vals'])
-vq = np.vectorize(weird_func)
-y = vq(x, order=0)
+vq = np.vectorize(lambda x: svm_func_obj(x, order=0, data=data))
+
+def pre_vectorized(x):
+    return  svm_func_obj(x, order=0, data=data)
+
+y = vq(pre_vectorized)
 plt.plot(x,y, 'o')
 plt.show()
 

### 2018-05-01 sqn/mnist/mnist_softmax.py
Auto stash before merge of "master" and "origin/master"

@@ -29,6 +29,9 @@ from tensorflow.examples.tutorials.mnist import input_data
 from tensorflow.contrib.opt.python.training.external_optimizer import ScipyOptimizerInterface
 from tensorflow.python.ops import variables
 
+sys.path.insert(0, '../optimizer/')
+from sqn_optimizer_hooks import SQNOptimizer
+
 import tensorflow as tf
 
 FLAGS = None
@@ -59,7 +62,7 @@ def main(_):
   cross_entropy = tf.reduce_mean(
       tf.nn.softmax_cross_entropy_with_logits(labels=y_, logits=y))
   #train_step = tf.train.GradientDescentOptimizer(0.5).minimize(cross_entropy)
-  optimizer = ScipyOptimizerInterface(loss=cross_entropy)
+  optimizer = SQNOptimizer(cross_entropy)
 
   with tf.Session() as sess:
     # Train

### 2018-05-01 sqn/optimizer/sd_lbfgs.py

@@ -84,7 +84,6 @@ class SdLBFGS():
 
 
     def sdlbfgs_step(self):
-        #print(f'{self._current_val}, {self._current_grad}')
         k = self._iterations
         p = self._mem_size
 
@@ -135,7 +134,6 @@ class SdLBFGS():
 
 
         self._mem_size = min(self._mem_size + 1, self._max_mem_size)
-#        print(f'(mem size: {self._mem_size})')
         if len(self._backward_errors) > self._max_mem_size:
             self._index_offset += 1
 

### 2018-05-01 sqn/optimizer/test.py

@@ -39,12 +39,9 @@ from hw4_functions import svm_objective_function as svm_func_obj
 x = np.array(result['iteration_vals'])
 vq = np.vectorize(lambda x: svm_func_obj(x, order=0, data=data))
 
-def pre_vectorized(x):
-    return  svm_func_obj(x, order=0, data=data)
-
-y = vq(pre_vectorized)
-plt.plot(x,y, 'o')
-plt.show()
+x = np.array(result['iteration_vals'])
+vq = np.vectorize(weird_func)
+y = vq(x, order=0)
 
 
 # vector = tf.Variable([7., 7.], 'vector')

### 2018-05-01 presentation/main.tex
More updates
Presentation

@@
-\documentclass[draft]{beamer}
+\documentclass{beamer}
 
@@
   \begin{itemize}
-    \item We implemented stochastic L-BFGS in \texttt{TensorFlow} following the
+    \item We implemented stochastic dampened (Sd) L-BFGS in \texttt{TensorFlow} following the
       work of Wang, Ma, Goldfarb, and Liu (2017).
@@
     \item Crucially, this holds even if the problem is \textbf{non-convex}.
-    \item Their update, \texttt{SdLBFGS}, 
-  \end{itemize}<++>
-
-\end{frame}<++>
+    \item Their update, \texttt{SdLBFGS}, preserves the positive-definiteness of
+      $H_k$ even in non-convex situations.
+  \end{itemize}
+\end{frame}
 
@@
 
+\begin{frame}[t]{Results: sanity checks}
+  \includegraphics[width=\textwidth]{../plots/sanity_checks.eps} 
+  \begin{itemize}
+    \item Since our code is in \texttt{NumPy}, we could test it by running
+      familiar homework functions.
+    \item Here all gradients are deterministic.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}[t]{Results: Higgs dataset}
+  \includegraphics[width=\textwidth]{../plots/higgs_dataset.eps} 
+\end{frame}
+
+\begin{frame}
+
+\end{frame}
+
 \end{document}

### 2018-05-06 paper/main.tex
Paper and presentation

\documentclass{article}

\usepackage[preprint]{nips_2018}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\title{A Stochastic Quasi-Newton Optimizer for TensorFlow}
\author{
  Jason Chen \\
  Department of Computer Science \\
  Stony Brook University \\
  Stony Brook, NY 11790 \\
  \texttt{hungrchen@cs.stonybrook.edu} \\
  \And
  David Kraemer \\
  Department of Applied Mathematics\\
  Stony Brook University \\
  Stony Brook, NY 11790 \\
  \texttt{davidkraemer@stonybrook.edu}
}

\date{May 6, 2018}

\begin{document}

\maketitle

\begin{abstract}
  We do some shit and it doesn't work.
\end{abstract}

\section{Introduction}

All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.

\section{Review of theory}

All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.

\section{Implementation}

All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.

\section{Empirical results}

All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.

\section{Discussion}

All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.

\end{document}

### 2018-05-06 presentation/main.tex
Paper and presentation

@@
   \begin{itemize}
-    \item Implementing L-BFGS proved challenging, and we discovered why it wasn't
+    \item Implementing SdLBFGS proved challenging, and we discovered why it wasn't
       part of the library already!
@@
       \href{https://github.com/tensorflow/tensorflow/issues/446}{Issue 446}, we
-      implemented L-BFGS as an \texttt{ExternalOptimizerInterface}.
+      implemented SdLBFGS as an \texttt{ExternalOptimizerInterface}.
     \item This allowed us to write \texttt{NumPy} code but ``plug in'' to

### 2018-05-06 paper/main.tex
Paper progress
Theory section is mostly done. This was the hardest part almost certainly.

@@
 
-\usepackage[preprint]{nips_2018}
+\usepackage[nonatbib,preprint]{nips_2018}
 
@@
 \usepackage{microtype}      % microtypography
+\usepackage{optidef}
+\usepackage{stmaryrd}
+\usepackage{mathtools}
+\usepackage{algpseudocode}
+\usepackage{algorithm}
+
+\usepackage[
+  backend=biber,
+  style=nature,
+  sorting=ynt
+]{biblatex}
+\addbibresource{main.bib}
+
+\newcommand{\RR}{\mathbb{R}}
+\DeclarePairedDelimiter{\norm}{\|}{\|}
+\DeclarePairedDelimiter{\ip}{\langle}{\rangle}
+\DeclarePairedDelimiter{\ex}{\mathbb{E}[}{]}
+\newcommand{\inv}{{}^{-1}}
+\newcommand{\trans}{{}^{\top}}
 
@@
 
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
+\begin{enumerate}
+  \item Introduce problem that SdLBFGS solves.
+  \item TensorFlow current state of the art.
+  \item Problems with the TensorFlow state.
+  \item Overview of paper.
+\end{enumerate}
 
@@
 
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
+\begin{enumerate}
+  \item Review main results of the Wang, Ma, Goldfarb, Liu paper
+    \cite{sdlbfgs}
+  \item State the SdLBFGS algortihm.
+  \item Mention some details about how the algorithm translates to code.
+\end{enumerate}
+
+The SdLBFGS algorithm is proposed by Wang, Ma, Goldfarb, and Liu \cite{sdlbfgs}
+for solving nonconvex stochastic optimization problems. In particular, it solves
+\begin{mini}
+  {x \in \RR^n}{f(x) = \ex{F(x, \xi)}}{}{} 
+  \label{prb:mini}
+\end{mini}
+for $F \in C^1(\RR^n \times \RR^d)$ and $\xi \in \RR^d$ is a random variable
+with cumulative distribution function $P$. Here $\ex{\cdot}$ denotes the
+mathematical expectation. In general $F$ need not be convex.
+
+The function $F$ is readily interpreted in the context of supervised statistical
+learning.  Given a corpus $\Xi$ of data, we put a distribution on $\Xi$ which
+reflects a sampling regime for a training set. The object of the learning
+process is, then, to the model parameter $x \in \RR^n$ which minimizes $F$
+\emph{in expectation} with respect to the training set. Provided that the
+sampling maintains some resemblance to the overall corpus, this gives a useful
+approximate parameter for the whole corpus.
+
+The SdLBFGS algorithm emerges from two concurrent optimization techniques.
+First, it follows in the tradition of approximate second-order algorithms which
+seek to perform a version of Newton's method while avoiding the computation of
+the Hessian $\nabla^2 f(x)$ explicitly. The classical BFGS approach is the
+canoncal quasi-Newton method. Second, it follows the development of stochastic
+algorithms which seek to randomly sample the components of the gradient $\nabla
+f(x)$ so that the essential convergence properties hold in expectation. In this
+sense it succeeds Stochastic Gradient Descent, among other canonical methods of
+this class. The limited memory constraint has antecedents in both streams of
+optimization techniques.
+
+Wang et al. \cite{sdlbfgs} assume that Problem (\ref{prb:mini}) satisfies the
+following assumptions:
+\begin{enumerate}
+  \item There is a real lower bound for $f$. That is, $\inf_{x \in \RR^n} f(x) >
+    -\infty$.
+  \item The gradient $\nabla f$ is Lipschitz continuous. That is, there exists
+    $L > 0$ such that 
+    \begin{equation*}
+      \norm{\nabla f(x) - \nabla f(y)} \leq L \norm{x - y}
+    \end{equation*}
+    for all $x,y \in \RR^n$. Moreover, $\nabla_{xx} F(x, \xi)$ exists and is
+    continuous and there exists $\kappa > 0$ with $\norm{\nabla^2_{xx} F(x,
+    \varepsilon)} \leq \kappa$ for any $x, \xi$.
+  \item The sampling regime is unbiased. That is, for the $k$th iterate the
+    following hold:
+    \begin{align*}
+      \ex{g(x_k, \xi_k)}_{\xi_k} &= \nabla f(x_k), \\
+      \ex{\norm{g(x_k, \xi_k) - \nabla f(x_k)}^2}_{\xi_k} &\leq \sigma^2,
+    \end{align*}
+    where $g(x, \xi_k) = \nabla F(x, \xi_k)$ and $\sigma^2$ is the noise level
+    of estimating $g(x, \xi_k)$. 
+\end{enumerate}
+
+For a convex function, $\nabla^2 f(x)$ is alwayas positive definite, but this
+need not be the case in for nonconvex functions. The work in Wang et al.
+\cite{sdlbfgs} shows how the BFGS approximate Hessian may be tweaked to ensure
+that positive definiteness is preserved even in nonconvex cases. The overall
+stochastic quasi-Newton method is described in Algorithm \ref{alg:sqn}, while
+the detailed implementation of SdLBFGS is given in Algorithm \ref{alg:sdlbfgs}.
+
+\begin{algorithm}
+  \begin{algorithmic}[1]
+    \Function{SQN}{$x_1, \ip{m_k}_{k=1}^{\infty},
+    \ip{\alpha_k}_{k=1}^{\infty}, p$}
+    \State{$\textrm{data} \leftarrow \emptyset$}
+    \For {$k = 1, 2, \dots$}
+    \State{$g_k \leftarrow \frac{1}{m_k} \sum_{i=1}^{m_k} g(x_k,
+    \varepsilon_{k,i})$} 
+    \State{$\Delta x_k, s_{k-1}, \bar{y}_{k-1}, \rho_{k-1} \leftarrow
+    \textsc{SdLBFGS}(g_k, {}^*(\textrm{data}\trans))$} \Comment{Tuple unpacking}
+    \State{$x_{k+1} \leftarrow x_k - \alpha_k \Delta x_k$}
+    \State{$\textit{append}(\textrm{data}, (s_{k-1}, \bar{y}_{k-1}, \rho_{k-1}))$}
+    \If{$k > p$}
+    \State{$\textit{pop}(\textrm{data})$}
+    \EndIf
+    \EndFor
+    \EndFunction
+  \end{algorithmic}
+  \label{alg:sqn}
+  \caption{%
+    The high level stochastic quasi-Newton minimization algorithm. Given an
+    initial point $x_1 \in \RR^n$, batch sizes $\ip{m_k}$, step sizes
+    $\ip{\alpha_k}$, and a maximum memory capacity $p$, perform BFGS with the
+    stochastic dampened direction update.
+  }
+\end{algorithm}
+
+\begin{algorithm}
+  \begin{algorithmic}[1]
+    \Function{SdLBFGS}{$g_{k-1}, \ip{s_{j}}_{j=k-p}^{k-2},
+    \ip{\bar{y}_{j}}_{j=k-p}^{k-2}, \ip{\rho_{j}}_{j=k-p}^{k-2}$}
+    \State{$\textrm{mem} \leftarrow \textit{min}(p, k-1)$}
+    \State{$s_{k-1} \leftarrow x_k - x_{k-1}$}
+    \State{$y_{k-1} \leftarrow \frac{1}{m_{k-1}}\sum_{i=1}^{m_{k-1}}[g(x_k,
+    \xi_{k-1,i}) - g(x_{k-1}, \xi_{k-1,i})]$}
+    \State{$\gamma_k \leftarrow \textit{max}\left( \frac{y_{k-1}\trans
+    y_{k-1}}{s\trans_{k-1}y_{k-1}}, \delta \right) \geq \delta$}
+    \State{%
+      $\theta \leftarrow
+      \textit{max}\left( 
+        \frac{3}{4} \frac{\gamma_{k}\inv s_{k-1}\trans
+        s_{k-1}}{\gamma_k\inv s_{k-1}\trans s_{k-1} - s_{k-1}\trans y_{k-1}}, 1
+      \right)
+    $}
+    \Comment{$\theta$ preserves positive definiteness.}
+    \State{$\bar{y}_{k-1} \leftarrow \theta y_{k-1} + (1 -\theta \gamma_{k}\inv s_{k-1})$}
+    \State{$\rho_{k-1} \leftarrow (s_{k-1}\trans \bar{y}_{k-1})\inv$}
+    \For{$i =0 , \dots, \textrm{mem}-1$}
+    \State{$\mu_i \leftarrow \rho_{k-i-1} u_i\trans s_{k-i-1}$} 
+    \State{$u_{i+1} \leftarrow u_i - \mu_i \bar{y}_{k-i-1}$}
+    \EndFor
+    \State{$v_0 \leftarrow \gamma_k\inv u_{\textrm{mem}}$}
+    \For{$i =0 , \dots, \textrm{mem}-1$}
+    \State{$\nu_i \leftarrow \rho_{k-\textrm{mem}+i} v_i\trans
+    \bar{y}_{k-\textrm{mem}+i}$}
+    \State{$v_{i+1} \leftarrow v_i + (\mu_{\textrm{mem}-i-1} - \nu_i)s_{k-\textrm{mem}+i} $}
+    \EndFor
+    \State{\Return{$v_p, s_{k-1}, \bar{y}_{k-1}, \rho_{k-1}$}}
+    \EndFunction
+  \end{algorithmic}
+  \caption{%
+    The SdLBFGS update step. The resulting output is $H_k g_k = v_p$,
+    where $H_k$ is the approximation of the $k$th iterate Hessian and $g_k$ is the
+    approximation of the $k$th iterate gradient. Note that the computation of $H_k$
+    is implicit, preventing additional storage requirements.
+  }
+  \label{alg:sdlbfgs}
+\end{algorithm}
 
@@
 
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
+\begin{enumerate}
+  \item Describe how TensorFlow algorithms are usually implemented.
+  \item Contrast with our implementation of SdLBFGS.
+  \item Explain some implementation issues.
+\end{enumerate}
 
@@
 
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
+\begin{enumerate}
+  \item Results on simple optimization problems. 
+  \item Results on potentially more difficult optimization problems.
+  \item Lack of results on MNIST with TensorFlow.
+\end{enumerate}
 
@@
 
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
-All work and no play makes Jack a dull boy.
+\begin{enumerate}
+  \item Explaining why the basic project --- implementing a useful practical
+    TensorFlow optimzier --- failed.
+  \item Takeaways from the results.
+  \item Future work.
+\end{enumerate}
+
+\printbibliography[heading=bibintoc]
 

### 2018-05-06 paper/main.tex
Introduction draft

@@
 
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
 \usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
 \usepackage{hyperref}       % hyperlinks
@@
 \title{A Stochastic Quasi-Newton Optimizer for TensorFlow}
+
 \author{
@@
 
-\begin{enumerate}
-  \item Introduce problem that SdLBFGS solves.
-  \item TensorFlow current state of the art.
-  \item Problems with the TensorFlow state.
-  \item Overview of paper.
-\end{enumerate}
+%\begin{enumerate}
+%  \item Introduce problem that SdLBFGS solves.
+%  \item TensorFlow current state of the art.
+%  \item Problems with the TensorFlow state.
+%  \item Overview of paper.
+%\end{enumerate}
+
+Limited memory variations of classical second-order minimization algorithms can
+prove highly attractive in applications with high-dimensional problems or when
+the cost of each iteration step is prohibitive under these classical
+formulations.  Specifically, the canonical second-order technique, Newton's
+method, involves explicit computation of the Hessian matrix of the objective
+function, whose time complexity grows with the square of the parameter
+dimension. Even with BFGS, which seeks to reduce the direct computational
+difficulty of exactly constructing the Hessian itself, encounters similar
+difficulties of scale. The limited memory approach is to bound this overall cost
+and ensure that approximate Hessians can be computed on existing hardware.
+
+We implement a variant of stochastic second-order algorithm as described by
+Wang, Ma, Goldfarb, and Liu \cite{sdlbfgs} for application in the
+\texttt{TensorFlow} library \cite{tensorflow}.  \texttt{TensorFlow} is a
+framework for graph-based computation models. Its primary use within the
+statistical learning community is in neural networks, which can be readily
+implemented by the library. After the model is constructed, the actual training
+is performed by an \texttt{Optimizer} object. In this setting, the
+\texttt{Optimizer} is presented with a partially observable objective function
+and gradient from which to perform optimization. Some typical \texttt{Optimizer}
+subclasses which are available in the Python API include
+an \texttt{GradientDescentOptimizer}\footnote{%
+  See
+  \href{https://github.com/tensorflow/tensorflow/blob/master/tensorflow/python/training/gradient_descent.py}{source.}
+}% 
+and an \texttt{AdagradOptimizer}\footnote{%
+  See \href{https://github.com/tensorflow/tensorflow/blob/master/tensorflow/python/training/adagrad.py}{source.}
+}%
+
+The presently available optimization algorithms implemented in the Python
+library are either first or zeroeth-order methods, requiring access to either
+the objective function or its gradient. The conspicuous absence of second-order
+methods, such as Newton or quasi-Newton algorithms, has been a subject of much
+discussion inside the \texttt{TensorFlow} community\footnote{%
+  The primary discussion has occurred surrounding
+  \href{https://github.com/tensorflow/tensorflow/issues/446}{Issue \#446}, in
+  which contributors have noted that the current infrastructure for developing
+  second-order methods is lacking and any implementation would require
+  considerable re-inventing the wheel.
+.}
+
+Following the suggestions of the community discussion, we present an
+implementation of Stochastic damped limited memory BFGS (SdLBFGS) for
+\texttt{TensorFlow} through the use of an \texttt{ExternalOptimizerInterface}.
+Presently used to wrap existing minimization routines implemented by outside
+libraries such as \texttt{SciPy}, it serves as a simple way to engage with the
+\texttt{TensorFlow} internals while avoiding a deep dive into its architecture.
+
+Our work proceeds as follows. Section 2 reviews the problem proposed in Wang et
+al and gives an explicit formulation of SdLBFGS according to their paper. We
+identify small modifications made to the core pseudocode and explain their
+necessity. Section 3 describes our SdLBFGS implementation and how it interfaces
+with \texttt{TensorFlow}. Section 4 shows preliminary experimental results on
+the correctness of our implementation on simple problems. Importantly, we report
+that our present implementation is incapable of giving meaningful progress on
+training practical models on common datasets. In Section 5, we discuss the
+implications of this negative result on future development and specify areas in
+which our project may be continued and improved.
+
+Source code for this project can be found at
+\url{https://github.com/DavidNKraemer/cse592-project}. Documentation, cruft
+removal, and user interface improvements are currently evolving.
 
@@
 
-For a convex function, $\nabla^2 f(x)$ is alwayas positive definite, but this
-need not be the case in for nonconvex functions. The work in Wang et al.
-\cite{sdlbfgs} shows how the BFGS approximate Hessian may be tweaked to ensure
-that positive definiteness is preserved even in nonconvex cases. The overall
-stochastic quasi-Newton method is described in Algorithm \ref{alg:sqn}, while
-the detailed implementation of SdLBFGS is given in Algorithm \ref{alg:sdlbfgs}.
+For a convex function, $\nabla^2 f(x)$ is always positive semidefinite (see Boyd
+and Vandenberghe \S 3.1.4 \cite{boyd}), but this need not be the case in for
+nonconvex functions. The work in Wang et al.  \cite{sdlbfgs} shows how the BFGS
+approximate Hessian may be tweaked to ensure that positive semidefiniteness is
+preserved even in nonconvex cases. The overall stochastic quasi-Newton method is
+described in Algorithm \ref{alg:sqn}, while the detailed implementation of
+SdLBFGS is given in Algorithm \ref{alg:sdlbfgs}.
 
@@
     $}
-    \Comment{$\theta$ preserves positive definiteness.}
+    \Comment{$\theta$ preserves positive semidefiniteness.}
     \State{$\bar{y}_{k-1} \leftarrow \theta y_{k-1} + (1 -\theta \gamma_{k}\inv s_{k-1})$}

### 2018-05-06 paper/main.tex
Section 2 draft

@@
     for all $x,y \in \RR^n$. Moreover, $\nabla_{xx} F(x, \xi)$ exists and is
-    continuous and there exists $\kappa > 0$ with $\norm{\nabla^2_{xx} F(x,
+    continuous, and there exists $\kappa > 0$ with $\norm{\nabla^2_{xx} F(x,
     \varepsilon)} \leq \kappa$ for any $x, \xi$.
@@
 
+We note that the formulation of Algorithm \ref{alg:sdlbfgs} differs slightly
+from Wang et al. In the original paper, the use of $\textrm{mem}$ in the section
+corresponding to lines 15 and 16 of Algorithm \ref{alg:sdlbfgs} is replaced by
+$p$. This leads to the difficulty that if $p$ indicates memory \emph{capacity}
+rather than simply the current amount of data in memory, the indices can become
+negative early in the iteration. Specifically, when $k < p$ in the original
+paper the algorithm has undefined behavior. We clarify this ambiguity in the
+following way: in our implementation, $p$ indicates the total memory capacity of
+the procedure, regardless of the total amount of memory presently in use. The
+quantity $\textrm{mem}$ replaces $p$ in the SdLBFGS update step so as to prevent
+array index out of bounds errors when $k$ is small (i.e., $k < p$), but both our
+implementation and the original paper's exhibit the same behavior whenever $k >
+p$.
+
 \begin{algorithm}
@@
     \ip{\alpha_k}_{k=1}^{\infty}, p$}
-    \State{$\textrm{data} \leftarrow \emptyset$}
+    \State{$\textrm{data} \leftarrow \textit{nil}()$}
     \For {$k = 1, 2, \dots$}

### 2018-05-06 paper/main.tex
Section 3 draft.

@@
 \usepackage{algorithm}
+\usepackage{listings}
+
+\usepackage{color}
+
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+
+\lstdefinestyle{mystyle}{
+  commentstyle=\color{codegreen},
+  keywordstyle=\color{magenta},
+  numberstyle=\tiny\color{codegray},
+  stringstyle=\color{codepurple},
+  basicstyle=\footnotesize\tt,
+  breakatwhitespace=false,         
+  breaklines=true,                 
+  captionpos=b,                    
+  keepspaces=true,                 
+  numbers=left,                    
+  numbersep=5pt,                  
+  showspaces=false,                
+  showstringspaces=false,
+  showtabs=false,                  
+  tabsize=2
+}
+
+\lstset{style=mystyle}
 
@@
 
+Our optimization object, \texttt{SQNOptimizer}, is implemented using
+\texttt{NumPy} \cite{numpy} with \texttt{ndarray} objects serving as the primary
+data structures for the numerical computations in Algorithm \ref{alg:sdlbfgs}.
+It inherits the features of the \texttt{ExternalOptimizerInterface}, which
+packages the objective function and its gradients from the network model in use
+and presents it in a \texttt{NumPy}-interoperable format.
+
+The general framework for using built-in \texttt{Optimizer} objects follows the
+structure presented in Figure \ref{fig:typical}.
+\begin{figure}[h]
+  \begin{lstlisting}[language=Python]
+# ... setup
+optimizer = Optimizer(**kwargs).minimize(objective_function)
+# ... setup
+with tf.Session() as session:
+    # ... training setup
+    optimizer.run(**training_kwargs)
+  \end{lstlisting}
+  \caption{The typical workflow for using built-in \texttt{Optimizer} objects in
+    a statistical learning setting. Here the \texttt{optimizer} is an operation
+    which is configured to minimize the objective function. The actual
+  minimization occurs inside of the session block.}
+  \label{fig:typical}
+\end{figure}
+The initialization of line 2 indicates that \texttt{optimizer} is a minimization
+operation, rather than the result of a minimization. By contrast, for subclasses
+of an \texttt{ExternalOptimizerInterface}, this structure is replaced by the
+workflow in Figure \ref{fig:external}.
+\begin{figure}[h]
+  \begin{lstlisting}[language=Python]
+# ... setup
+optimizer = ExternalOptimizer(objective_function, **kwargs)
+# ... setup
+with tf.Session() as session:
+    # ... training setup
+    optimizer.minimize(session, **training_kwargs)
+  \end{lstlisting}
+  \caption{%
+    The workflow for using objects of 
+  \texttt{ExternalOptimizerInterface} subclasses, such as the implementation of
+  \texttt{SQNOptimizer}. Here the \texttt{optimizer} itself is the minimizer while
+  the \texttt{minimize} method actually performs estimation of the minimal
+  objective function value.  }
+  \label{fig:external}
+\end{figure}
+
+Now, the \texttt{minimize} method returns the \emph{result} of a minimization
+rather than a minimization operation. As \texttt{SQNOptimizer} inherits this
+functionality, the second workflow is required.
+
 \section{Empirical results}

### 2018-05-06 sqn/optimizer/algorithms.py
Fix Numerical Problem and add plot script

@@ -4,46 +4,40 @@
 
 import time
 import numpy as np
+import math
 
 
-def bisection( one_d_fun, MIN, MAX, eps=1e-5, maximum_iterations=65536 ):
-    """
-    Identify a critical point of a single-variable real-valued function on a
-    specified interval. 
-    one_d_fun:          a tuple of the form (f, f'), where f : R -> R is differentiable 
-    MIN:                lower bound on the interval on which searching occurs 
-    MAX:                upper bound on the interval on which searching occurs 
-    eps:                the maximum allowed error in the resulting stepsize t
-    maximum_iterations: the maximum allowed number of iterations
-    """
+def bisection( one_d_fun, MIN, MAX, eps=1e-8, maximum_iterations=65536 ):
 
-    # counting the number of iterations
-    iterations = 0
+  # counting the number of iterations
+  iterations = 0
 
-    if eps <= 0:
-        raise ValueError("Epsilon must be positive")
+  if eps <= 0:
+      raise ValueError("Epsilon must be positive")
 
-    while True:
+  while True:
 
-        iterations += 1
+    MID = ( MAX + MIN ) / 2
 
-        MID = ( MAX + MIN ) / 2
+    # Oracle access to the function value and derivative
+    value, derivative = one_d_fun( MID, 1 )
 
-        # Oracle access to the function value and derivative
-        value, derivative = one_d_fun( MID, 1 )
+    # f(MID)-f(x^*) <= |f'(MID)*(MID-x^*)| <= |f'(MID)|*(MAX-MID)
+    suboptimality = abs(derivative) * (MAX - MID)
 
-        if np.abs(derivative) < eps or (MAX - MIN) < 2 * eps:
-            break
-        elif derivative < 0.:
-            MIN = MID
-        else:
-            MAX = MID
+    if suboptimality <= eps:
+        break
 
-        iterations += 1
-        if iterations>=maximum_iterations:
-            break
+    if derivative > 0:
+        MAX=MID
+    else:
+        MIN=MID
 
-    return MID
+    iterations += 1
+    if iterations>=maximum_iterations:
+        break
+
+  return MID
 
 
 def function_on_line(func, x, direction, order):
@@ -62,7 +56,7 @@ def function_on_line(func, x, direction, order):
 
 ###############################################################################
 def exact_line_search( func, x, direction, eps=1e-9, maximum_iterations=65536 ):
-    """ 
+    """
     'Exact' linesearch (using bisection method)
[…]

### 2018-05-06 sqn/optimizer/hw1_functions.py

@@ -10,8 +10,9 @@ from math import exp
 def weird_func( x, order=0 ):
 
   # f(x) = x^4 + 6x^2 + 12(x-4)e^(x-1)
+#  print(x)
   value = pow(x, 4) + 6 * pow(x, 2) + 12 * (x - 4) * exp(x - 1)
-  
+
   if order==0:
       return value
   elif order==1:
@@ -30,13 +31,61 @@ def weird_func( x, order=0 ):
   else:
         raise ValueError("The argument \"order\" should be 0, 1 or 2")
 
+
+########################################################333
+def no_conv_func1(x, order=0):
+  '''x^4-4x^3-x^2+10x'''
+#  print(x)
+  value = pow(x, 4) - 4*pow(x, 3) - pow(x,2) + 10*x
+
+  if order==0:
+      return value
+  elif order==1:
+      # f'(x) = 4x^3 + 12x + 12(x-3)e^(x-1)
+      gradient = 4*pow(x, 3) - 12*pow(x, 2) - 2*x + 10
+
+      return (value, gradient)
+  elif order==2:
+      # f'(x) = 4x^3 + 12x + 12(x-3)e^(x-1)
+      gradient = 4*pow(x, 3) - 12*pow(x, 2) - 2*x + 10
+
+      # f''(x)= 12 (1 + e^(-1 + x) (-2 + x) + x^2)
+      hessian = 12*pow(x, 3) - 24*x - 2
+
+      return (value, gradient, hessian)
+  else:
+        raise ValueError("The argument \"order\" should be 0, 1 or 2")
+
+def no_conv_func2(x, order=0):
+  ''''''
+#  print(x)
+  value = pow(x, 6) - 2*pow(x, 5) + pow(x,3) - pow(x,2) + 3*x
+
+  if order==0:
+      return value
+  elif order==1:
+      # f'(x) = 4x^3 + 12x + 12(x-3)e^(x-1)
+      gradient = 6 * pow(x, 5) - 10*pow(x, 4) + 3*pow(x,2) - 2*x + 3
+
+      return (value, gradient)
+  elif order==2:
+      # f'(x) = 4x^3 + 12x + 12(x-3)e^(x-1)
+      gradient = 6 * pow(x, 5) - 10*pow(x, 4) + 3*pow(x,2) - 2*x + 3
+
+      # f''(x)= 12 (1 + e^(-1 + x) (-2 + x) + x^2)
[…]

### 2018-05-06 sqn/optimizer/sd_lbfgs.py

@@ -5,21 +5,28 @@ Created on Thu Apr 19 02:07:39 2018
 @author: vbigmouse
 """
 from numpy import dot, zeros, asarray, sqrt
+from math import isnan, isinf
 from scipy.optimize import OptimizeResult
+from time import time
+from pprint import pprint
 
-
-def harmonic_sequence(k):
+def harmonic_sequence(k, init_step_size):
     # For any nonnegative integer
-    return 0.9/(sqrt(k+1))
+    return init_step_size /(k+1)
+#    return 10 / (k+1)
+def sqrt_sequence(k, init_step_size):
+    # For any nonnegative integer
+    return init_step_size /(sqrt(k+1))
 
 class SdLBFGS():
     def __init__(self, func, initial_val, *args,
             max_iterations=1000,
             mem_size=10,
             batch_size=50,
+            init_step_size=0.1,
             step_size=harmonic_sequence,
-            delta=0.01,
-            tol=1e-4,
+            delta=0.1,
+            tol=1e-6,
             **kwargs):
 
         self._func = func
@@ -29,7 +36,8 @@ class SdLBFGS():
         self._max_mem_size = mem_size
         self._mem_size = 0
         self._batch_size = batch_size
-        self._step_size = step_size
+        self._step_size = lambda k:step_size(k, self._init_step_size)
+        self._init_step_size = init_step_size
         self._delta = delta
         self._tolerance = tol
 
@@ -38,12 +46,18 @@ class SdLBFGS():
         self._previous_val = zeros(self._initial_val.shape)
         self._current_grad = zeros(self._initial_val.shape)
         self._previous_grad = zeros(self._initial_val.shape)
+        self._current_objval = asarray(self._func(self._current_val)[0])
 
         self._backward_errors = ShiftList(self._max_mem_size)
         self._ybars = ShiftList(self._max_mem_size)
         self._rhos = ShiftList(self._max_mem_size)
 
         self._iteration_vals = [self._current_val]
+        self._iteration_grads = [self._current_grad]
+        self._iteration_objvals = [self._current_objval]
+
+        self._start_time = time()
+        self._iteration_runtimes = [0]
 
         self._result = OptimizeResult()
         self._result['success'] = False
@@ -52,22 +66,37 @@ class SdLBFGS():
 
     def run(self):
         while not self._result['success'] and self._iterations < self._max_iterations:
+#            pprint(self.__dict__, indent=2)
             self.sqn_step()
[…]

### 2018-05-06 sqn/optimizer/test.py

@@ -1,8 +1,11 @@
-#import tensorflow as tf
+import tensorflow as tf
 from sqn_optimizer_hooks import SQNOptimizer
 from sd_lbfgs import SdLBFGS
-from hw1_functions import weird_func
+from sd_lbfgs import harmonic_sequence as harmonic_seq
+from sd_lbfgs import sqrt_sequence as sqrt_seq
 from hw4_functions import svm_objective_function_stochastic as svm_func
+from hw4_functions import svm_objective_function as svm_func_obj
+import matplotlib.pyplot as plt
 import numpy as np
 from time import sleep
 from pprint import pprint
@@ -12,51 +15,541 @@ def quadratic(x, order=1):
         return x * x
     if order==1:
         return x * x, 2 * x
+def norm_z(x):
+    x=(x-x.mean())/x.std()
+#    m=np.median(x)
+#    x.fillna(m,inplace=True)
+    return x
+
+def norm_m(x):
+    x = (x - x.min() + 1) / (x.max() - x.min())
+#    m = np.median(x)
+#    x.fillna(m, inplace=True)
+    return x
 
 #%%
-data = np.loadtxt('HIGGS_subset.csv', delimiter=',')
+rawdata = np.loadtxt('HIGGS_subset.csv', delimiter=',')
 
-labels = np.asmatrix(2*data[:,0]-1).T
-features = np.asmatrix(data[:,1:])
+labels = np.asmatrix(2 * rawdata[:, 0] -1).T
+features = np.asmatrix(rawdata[:, 1:])
+features = norm_z(features)
+
 data = np.concatenate((features, labels), axis = 1)
 d = features.shape[1]
 w = np.zeros((d,1))
 #%%
-
-#optimizer = SdLBFGS(lambda x: weird_func(x, order=1), 1., batch_size=2)
-
-
 optimizer = SdLBFGS(lambda x: svm_func(x, order=1, data=data, minibatch_size=50), w,
         batch_size=50)
 
-# for i in range(10):
-#     optimizer.sqn_step()
+result = optimizer.run()
+#%%
+'''Logistic Resression test'''
+
+'''General Setting'''
+max_iterations=1000
+points_to_plot=1000
+
+from sd_lbfgs import SdLBFGS
+from hw4_functions import cross_entropy_error as cross_enp_func
+
+#%%
[…]

### 2018-05-06 sqn/optimizer/hw4_functions.py
Add some object functions

@@ -1,0 +1,171 @@
+######
+######  This file includes different functions used in HW3
+######
+
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+
+import random
+import math
+
+
+
+def svm_objective_function(w, data, order):
+    labels = data[:,-1]
+    features = data[:,:-1]
+
+    n=len(labels)
+
+    subgradient = np.zeros(w.shape)
+#    result = []
+    i = 0
+#    sample_ind = random.sample(range(0, n), 10000)
+    if order==0:
+#        for w_v in w.T:
+#            if i % 10 == 0:
+        value = 0
+        for i in range(0, n):
+            value += int(max(1 - np.dot(labels[i], np.dot(features[i], w)), 0))
+#        result.append(value)
+        return value
+#        return result
+    elif order==1:
+        for i in range(0, n):
+            if labels[i].item(0) * features[i] * w < 1:
+                subgradient += -np.multiply(labels[i], features[i].T) / n
+            value += max(1 - labels[i] * np.dot(features[i].T, w), 0)
+        return (value, subgradient)
+    else:
+        raise ValueError("The argument \"order\" should be 0 or 1")
+
+def svm_objective_function_stochastic(w, data, order, minibatch_size):
+
+    labels = data[:,-1]
+    features = data[:,:-1]
+    n = len(labels)
+    value = 0
+    subgradient = np.zeros(w.shape)
+    sample_ind = random.sample(range(0, n), minibatch_size)
+
+    if order==0:
+        for i in sample_ind:
+            value += max(1 - labels[i] * features[i] * w , 0)
+        return value
+    elif order==1:
+        for i in sample_ind:
+            if labels[i].item(0) * features[i] * w < 1:
+                subgradient += -np.multiply(labels[i], features[i].T) / minibatch_size
+            value += max(1 - labels[i] * features[i] * w, 0)
+
+
+        return (value, subgradient)
+    else:
+        raise ValueError("The argument \"order\" should be 0 or 1")
+
+################################################################
+
+
+def logrithm_loss(w, data, order, minibatch_size):
+    labels = data[:,-1]
+    features = data[:,:-1]
+    n=len(labels)
+    value = 0
+    subgradient = np.zeros(w.shape)
+
+    sample_ind = random.sample(range(0, n), minibatch_size)
+
+    if order==0:
+        for i in sample_ind:
+            x = features[i].T
[…]

### 2018-05-06 sqn/optimizer/nonconvex_x6_script.py
Script for nonconvex_x6 function opt and plot

@@ -1,0 +1,163 @@
+# -*- coding: utf-8 -*-
+from sd_lbfgs import SdLBFGS
+from sd_lbfgs import harmonic_sequence as harmonic_seq
+from sd_lbfgs import sqrt_sequence as sqrt_seq
+import matplotlib.pyplot as plt
+import numpy as np
+from sd_lbfgs import SdLBFGS
+
+
+#%%
+'''non-convex function'''
+
+points_to_plot=100
+max_iterations = 10000
+
+from hw1_functions import no_conv_func2
+obj_f = lambda x : no_conv_func2(x, order=1)
+
+
+initial_x = 2.
+
+optimizer = SdLBFGS(obj_f, initial_x,
+                    batch_size=1,
+                    delta=0.01,
+                    step_size=sqrt_seq,
+#                    step_size=harmonic_seq,
+                    mem_size=20,
+                    init_step_size=0.1,
+                    max_iterations=10000)
+
+result = optimizer.run()
+
+sdlbfs_x =result['iteration_vals']
+sdlbfs_values =result['iteration_objvals']
+sdlbfs_runtimes =result['iteration_runtimes']
+
+
+print('Solution found by sdLBFGS', sdlbfs_x[-1])
+print('Objective function', no_conv_func2(sdlbfs_x[-1],0))
+
+sdlbfs_x =sdlbfs_x[0::int(len(sdlbfs_x)/min(len(sdlbfs_runtimes), points_to_plot))]
+sdlbfs_values =sdlbfs_values[0::int(len(sdlbfs_values)/min(len(sdlbfs_runtimes), points_to_plot))]
+sdlbfs_runtimes =sdlbfs_runtimes[0::int(len(sdlbfs_runtimes)/min(len(sdlbfs_runtimes), points_to_plot))]
+
+'SGD'
+import algorithms as alg
+
+obj_f = lambda x, order: no_conv_func2(x, order)
+initial_x = 2.
+sgd_x, sgd_values, sgd_runtimes, sgd_xs = \
+    alg.subgradient_descent(obj_f, initial_x, max_iterations, 0.01)
+
+print('Solution found by stochastic subgradient descent', sgd_x)
+print('Objective function', obj_f(sgd_x,0))
+sgd_its = len(sgd_runtimes)
[…]

### 2018-05-06 mnist/mnist_fully_connected.py
Auto stash before merge of "master" and "origin/master"

@@ -38,9 +38,12 @@ import numpy as np
 
 sys.path.insert(0, '../optimizer/')
 
+from sqn_optimizer_hooks import SQNOptimizer
 
-from external_optimizer import ScipyOptimizerInterface
-Optimizer = ScipyOptimizerInterface
+
+Optimizer = SQNOptimizer
+memsize = 2
+batch_size = 10
 
 FLAGS = None
 
@@ -80,7 +83,8 @@ def main(_):
     out = tf.matmul(h_fc2, W_fc3) + b_fc3
 
     cross_entropy = tf.reduce_mean(tf.nn.softmax_cross_entropy_with_logits(labels=y_, logits=out))
-    optimizer = Optimizer(cross_entropy, options={'maxiter': 1})
+    optimizer = Optimizer(cross_entropy, mem_size=memsize, batch_size=batch_size,
+            max_iterations=50)
     correct_prediction = tf.equal(tf.argmax(out, 1), tf.argmax(y_, 1))
     accuracy = tf.reduce_mean(tf.cast(correct_prediction, tf.float32))
 
@@ -89,7 +93,7 @@ def main(_):
         for i in range(600*40):
             print(f"{i}", end="\r")
             batch = mnist.train.next_batch(100)
-            if i % 200 == 0:
+            if i % 600 == 0:
                 test_batch_size = 10000
                 batch_num = int(mnist.train.num_examples / test_batch_size)
                 train_loss = 0
@@ -107,6 +111,8 @@ def main(_):
                 test_err = 1-accuracy.eval(feed_dict={x: mnist.test.images, y_: mnist.test.labels})
 
                 print('epoch %d, training cost %g, test error %g ' % (i/600, train_loss, test_err))
+                with open(f'{memsize}_memory_{batch_size}_batch.csv', 'a') as f:
+                    f.write(f'{i/600},{train_loss},{test_err}\n')
             optimizer.minimize(sess, feed_dict={x: batch[0], y_: batch[1]})
 
 

### 2018-05-06 mnist/mnist_softmax.py

@@ -62,11 +62,13 @@ def main(_):
   cross_entropy = tf.reduce_mean(
       tf.nn.softmax_cross_entropy_with_logits(labels=y_, logits=y))
   #train_step = tf.train.GradientDescentOptimizer(0.5).minimize(cross_entropy)
-  optimizer = SQNOptimizer(cross_entropy)
+  optimizer = SQNOptimizer(cross_entropy, batch_size=10, mem_size=1,
+          max_iterations=100)
 
   with tf.Session() as sess:
     # Train
-    for _ in range(1000):
+    for i in range(1000):
+      print(i,end='\r')  
       batch_xs, batch_ys = mnist.train.next_batch(100)
       sess.run(fetches=tf.global_variables_initializer())
       optimizer.minimize(sess, feed_dict={x: batch_xs, y_: batch_ys})

### 2018-05-06 optimizer/sd_lbfgs.py

@@ -4,12 +4,17 @@ Created on Thu Apr 19 02:07:39 2018
 
 @author: vbigmouse
 """
+<<<<<<< Updated upstream:sqn/optimizer/sd_lbfgs.py
 from numpy import dot, zeros, asarray, sqrt
 from math import isnan, isinf
+=======
+from numpy import dot, zeros, asarray, sqrt, power
+>>>>>>> Stashed changes:optimizer/sd_lbfgs.py
 from scipy.optimize import OptimizeResult
 from time import time
 from pprint import pprint
 
+<<<<<<< Updated upstream:sqn/optimizer/sd_lbfgs.py
 def harmonic_sequence(k, init_step_size):
     # For any nonnegative integer
     return init_step_size /(k+1)
@@ -17,16 +22,28 @@ def harmonic_sequence(k, init_step_size):
 def sqrt_sequence(k, init_step_size):
     # For any nonnegative integer
     return init_step_size /(sqrt(k+1))
+=======
+
+def default_step_size(k):
+    # For any nonnegative integer
+    return power(k+1,-0.5)
+>>>>>>> Stashed changes:optimizer/sd_lbfgs.py
 
 class SdLBFGS():
     def __init__(self, func, initial_val, *args,
             max_iterations=1000,
             mem_size=10,
             batch_size=50,
+<<<<<<< Updated upstream:sqn/optimizer/sd_lbfgs.py
             init_step_size=0.1,
             step_size=harmonic_sequence,
             delta=0.1,
             tol=1e-6,
+=======
+            step_size=default_step_size,
+            delta=0.01,
+            tol=1e-4,
+>>>>>>> Stashed changes:optimizer/sd_lbfgs.py
             **kwargs):
 
         self._func = func
@@ -52,12 +69,17 @@ class SdLBFGS():
         self._ybars = ShiftList(self._max_mem_size)
         self._rhos = ShiftList(self._max_mem_size)
 
+<<<<<<< Updated upstream:sqn/optimizer/sd_lbfgs.py
         self._iteration_vals = [self._current_val]
         self._iteration_grads = [self._current_grad]
         self._iteration_objvals = [self._current_objval]
 
         self._start_time = time()
         self._iteration_runtimes = [0]
+=======
+        self._iteration_vals = []
+        self._iteration_grad_errs = []
+>>>>>>> Stashed changes:optimizer/sd_lbfgs.py
 
         self._result = OptimizeResult()
[…]

### 2018-05-06 sqn.py

@@ -1,88 +1,0 @@
-import numpy as np
-
-def sqn(x0, H1, m, alpha, g, dist):
-    """
-    Stochastic Quasi Newton's Method
-    x0:    the starting point
-    H1:    the initial approximation of the gradient
-    m:     a sequence of batch sizes
-    alpha: a sequence of step sizes
-    g:     R^n x R^d -> R, stochastic gradient of f taking location and random input
-    dist:  the distribution of the random variable
-
-    """
-
-    # iteration parameters
-    MAX_ITER = 100
-    tol = 1e-8
-    p = 4  # memory size, don't know what this means
-
-
-    # iterate data, which is computed and eventually to be returned     
-    x = np.empty((MAX_ITER+1, *x0.shape))
-    s = np.empty((MAX_ITER+1, *x0.shape))
-    y = np.empty((MAX_ITER+1, *x0.shape))
-    ybar = np.empty((MAX_ITER+1, *x0.shape))
-    err = np.empty(MAX_ITER+1)
-    rho = np.empty(MAX_ITER+1)
-
-    x[0] = x0
-    err[0] = np.linalg.norm(x0)
-
-    k = 0
-
-    # main loop
-    while err[k] >= tol and k < MAX_ITER:
-        # general SQN step
-        x[k+1] = x[k] - alpha[k] * sdlbfgs_step(k, dist, m, g, x, s, y, p, rho, ybar)
-        err[k+1] = np.linalg.norm(x[k+1] - x[k])
-
-    return x[:k], err[:k]
-
-
-def sdlbfgs_step(k, dist, m, g, x, s, y, p, rho, ybar):
-    """
-    Step computation using SdLBFGS
-    k:    current iteration
-    dist: the distribution of the random variable
-    m:    sequence of batch sizes
-    g:    R^n x R^d -> R, stochastic gradient of f taking location and random input
-    x:    sequence of points in algorithm [TODO: this needs rewriting]
-    s:    blah 
-    y:    blah
-    """
-
-    sub_iter_len = min(p, k-1)
-    mu = np.empty((sub_iter_len, 1))
-    u = np.empty((sub_iter_len, *y[0].shape))
-    u[0] = g[k]
-
[…]

### 2018-05-06 optimizer/algorithms.py
Modify some functions

@@ -125,7 +125,6 @@ def backtracking_line_search( func, x, direction, alpha=0.4, beta=0.9, maximum_i
     t = 1
     iterations = 0
     while True:
-
         if func(x+t*direction, 0)<value_0 + alpha * t * gradient_0.T*direction:
             break
 
@@ -268,6 +267,7 @@ def bfgs( func, initial_x, initial_inv_h, eps=1e-6, maximum_iterations=65536, li
     values = []
     runtimes = []
     xs = []
+    grads = []
     start_time = time.time()
     m = len( x )
     iterations = 0
@@ -286,7 +286,7 @@ def bfgs( func, initial_x, initial_inv_h, eps=1e-6, maximum_iterations=65536, li
         values.append( value )
         runtimes.append( time.time() - start_time )
         xs.append( x.copy() )
-
+        grads.append(gradient.copy())
         # termination criterion
         if np.vdot( gradient, gradient ) <= eps:
             break
@@ -313,7 +313,7 @@ def bfgs( func, initial_x, initial_inv_h, eps=1e-6, maximum_iterations=65536, li
         if iterations >= maximum_iterations:
             break
 
-    return (x, values, runtimes, xs)
+    return (x, values, runtimes, xs, grads)
 
 ###############################################################################
 def objective_log_barrier( f, phi, x, t, order=0 ):
@@ -473,8 +473,10 @@ def subgradient_descent( func, initial_x, maximum_iterations=65536, initial_step
     values = []
     runtimes = []
     xs = []
+    grads = []
     start_time = time.time()
     iterations = 0
+
 #    stepsize = initial_stepsize
 
     # subgradient updates
@@ -489,7 +491,7 @@ def subgradient_descent( func, initial_x, maximum_iterations=65536, initial_step
         values.append( value )
         runtimes.append( time.time() - start_time )
         xs.append( x.copy() )
-
+        grads.append(gradient.copy())
         # x = ( TODO: update of subgradient descent )
         if iterations:
             x -= initial_stepsize / np.sqrt(iterations) * gradient
@@ -499,7 +501,7 @@ def subgradient_descent( func, initial_x, maximum_iterations=65536, initial_step
         if iterations >= maximum_iterations:
             break
 
-    return (x, values, runtimes, xs)
+    return (x, values, runtimes, xs, grads)
 
 
[…]

### 2018-05-06 sqn/optimizer/hw4_functions.py

@@ -138,7 +138,7 @@ def cross_entropy_error(w, data, order, minibatch_size):
     else:
         raise ValueError("The argument \"order\" should be 0 or 1")
 
-##################################################################3
+###################################################################
 def tanh_regul_error(w, data, order, minibatch_size, regularizer=1e-4):
     labels = data[:,-1]
     features = data[:,:-1]

### 2018-05-06 sqn/optimizer/nonconvex_x6_script.py

@@ -4,14 +4,14 @@ from sd_lbfgs import harmonic_sequence as harmonic_seq
 from sd_lbfgs import sqrt_sequence as sqrt_seq
 import matplotlib.pyplot as plt
 import numpy as np
-from sd_lbfgs import SdLBFGS
+
 
 
 #%%
 '''non-convex function'''
 
 points_to_plot=100
-max_iterations = 10000
+max_iterations = 1000
 
 from hw1_functions import no_conv_func2
 obj_f = lambda x : no_conv_func2(x, order=1)
@@ -26,28 +26,30 @@ optimizer = SdLBFGS(obj_f, initial_x,
 #                    step_size=harmonic_seq,
                     mem_size=20,
                     init_step_size=0.1,
-                    max_iterations=10000)
+                    max_iterations=1000)
 
 result = optimizer.run()
 
 sdlbfs_x =result['iteration_vals']
 sdlbfs_values =result['iteration_objvals']
 sdlbfs_runtimes =result['iteration_runtimes']
-
+sdlbfs_grads = result['iteration_grads']
 
 print('Solution found by sdLBFGS', sdlbfs_x[-1])
 print('Objective function', no_conv_func2(sdlbfs_x[-1],0))
 
-sdlbfs_x =sdlbfs_x[0::int(len(sdlbfs_x)/min(len(sdlbfs_runtimes), points_to_plot))]
-sdlbfs_values =sdlbfs_values[0::int(len(sdlbfs_values)/min(len(sdlbfs_runtimes), points_to_plot))]
-sdlbfs_runtimes =sdlbfs_runtimes[0::int(len(sdlbfs_runtimes)/min(len(sdlbfs_runtimes), points_to_plot))]
+sdlbfs_its = len(sdlbfs_runtimes)
+sdlbfs_x =sdlbfs_x[0::int(sdlbfs_its/min(sdlbfs_its, points_to_plot))]
+sdlbfs_values =sdlbfs_values[0::int(sdlbfs_its/min(sdlbfs_its, points_to_plot))]
+sdlbfs_runtimes =sdlbfs_runtimes[0::int(sdlbfs_its/min(sdlbfs_its, points_to_plot))]
+sdlbfs_grads = sdlbfs_grads[0::int(sdlbfs_its/min(sdlbfs_its, points_to_plot))]
 
 'SGD'
 import algorithms as alg
 
[…]

### 2018-05-06 optimizer/sd_lbfgs.py
Fix some conflict message.

@@ -4,17 +4,14 @@ Created on Thu Apr 19 02:07:39 2018
 
 @author: vbigmouse
 """
-<<<<<<< Updated upstream:sqn/optimizer/sd_lbfgs.py
-from numpy import dot, zeros, asarray, sqrt
-from math import isnan, isinf
-=======
+
 from numpy import dot, zeros, asarray, sqrt, power
->>>>>>> Stashed changes:optimizer/sd_lbfgs.py
+from math import isnan, isinf
 from scipy.optimize import OptimizeResult
 from time import time
 from pprint import pprint
 
-<<<<<<< Updated upstream:sqn/optimizer/sd_lbfgs.py
+
 def harmonic_sequence(k, init_step_size):
     # For any nonnegative integer
     return init_step_size /(k+1)
@@ -22,28 +19,23 @@ def harmonic_sequence(k, init_step_size):
 def sqrt_sequence(k, init_step_size):
     # For any nonnegative integer
     return init_step_size /(sqrt(k+1))
-=======
+
 
 def default_step_size(k):
     # For any nonnegative integer
     return power(k+1,-0.5)
->>>>>>> Stashed changes:optimizer/sd_lbfgs.py
+
 
 class SdLBFGS():
     def __init__(self, func, initial_val, *args,
             max_iterations=1000,
             mem_size=10,
             batch_size=50,
-<<<<<<< Updated upstream:sqn/optimizer/sd_lbfgs.py
             init_step_size=0.1,
-            step_size=harmonic_sequence,
-            delta=0.1,
-            tol=1e-6,
-=======
             step_size=default_step_size,
             delta=0.01,
             tol=1e-4,
->>>>>>> Stashed changes:optimizer/sd_lbfgs.py
+
             **kwargs):
 
         self._func = func
@@ -69,17 +61,15 @@ class SdLBFGS():
         self._ybars = ShiftList(self._max_mem_size)
         self._rhos = ShiftList(self._max_mem_size)
 
-<<<<<<< Updated upstream:sqn/optimizer/sd_lbfgs.py
-        self._iteration_vals = [self._current_val]
-        self._iteration_grads = [self._current_grad]
-        self._iteration_objvals = [self._current_objval]
 
-        self._start_time = time()
-        self._iteration_runtimes = [0]
-=======
         self._iteration_vals = []
+        self._iteration_grads = []
+        self._iteration_objvals = []
+
         self._iteration_grad_errs = []
->>>>>>> Stashed changes:optimizer/sd_lbfgs.py
[…]

### 2018-05-07 paper/main.tex
Auto stash before merge of "master" and "origin/master"
Plotting rewrites and inclusion in paper
I rewrote some of the nonconvex function plot script so that it would fit better in the paper. I also added the two plots I have right now to the paper.

@@
 \usepackage{optidef}
+\usepackage[]{graphicx}
 \usepackage{stmaryrd}
@@
 \usepackage{listings}
+\usepackage[outdir=../plots/]{epstopdf}
 
@@
 \DeclarePairedDelimiter{\ip}{\langle}{\rangle}
+\DeclarePairedDelimiter{\set}{\{}{\}}
 \DeclarePairedDelimiter{\ex}{\mathbb{E}[}{]}
@@
 
-\begin{enumerate}
-  \item Describe how TensorFlow algorithms are usually implemented.
-  \item Contrast with our implementation of SdLBFGS.
-  \item Explain some implementation issues.
-\end{enumerate}
+% \begin{enumerate}
+%   \item Describe how TensorFlow algorithms are usually implemented.
+%   \item Contrast with our implementation of SdLBFGS.
+%   \item Explain some implementation issues.
+% \end{enumerate}
 
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\textwidth]{../plots/nonconvex_results}
+  \caption{
+    SdLBFGS compared with traditional BFGS together with AdaGrad and Stochastic
+    Gradient Descent on the problem $f(x) = x^6 - 2x^5 + x^3 - x^2 + 3x$, shown
+    on the right. Each algorithm is initialized near the local minimum at around
+    $x = 1.6$. We see that SGD, AdaGrad, and BFGS quickly converge to the local
+    minimum, whereas SdLBFGS converges to the global minimum at around $x =
+    -0.8$ more slowly.
+  }
+  \label{fig:convex-linear-system}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\textwidth]{../plots/linear_system_loss}
+  \caption{
+    SdLBFGS performance on the solving the overdetermined linear system
+    $Ax = b$ with $A \in \RR^{50 \times 100}$ with convex smooth loss function
+    $\norm{Ax-b}_2^2$. We sweep different memory capacity values $p \in
+    \set{1,10,20,30}$ and observe that convergence is approximately equivalent
+    between configurations.
+  }
+  \label{fig:convex-linear-system}
+\end{figure}
+
 \section{Discussion}

### 2018-05-07 optimizer/sd_lbfgs.py

@@ -4,17 +4,12 @@ Created on Thu Apr 19 02:07:39 2018
 
 @author: vbigmouse
 """
-<<<<<<< Updated upstream:sqn/optimizer/sd_lbfgs.py
-from numpy import dot, zeros, asarray, sqrt
-from math import isnan, isinf
-=======
 from numpy import dot, zeros, asarray, sqrt, power
->>>>>>> Stashed changes:optimizer/sd_lbfgs.py
+from math import isnan, isinf
 from scipy.optimize import OptimizeResult
 from time import time
 from pprint import pprint
 
-<<<<<<< Updated upstream:sqn/optimizer/sd_lbfgs.py
 def harmonic_sequence(k, init_step_size):
     # For any nonnegative integer
     return init_step_size /(k+1)
@@ -22,28 +17,21 @@ def harmonic_sequence(k, init_step_size):
 def sqrt_sequence(k, init_step_size):
     # For any nonnegative integer
     return init_step_size /(sqrt(k+1))
-=======
 
 def default_step_size(k):
     # For any nonnegative integer
     return power(k+1,-0.5)
->>>>>>> Stashed changes:optimizer/sd_lbfgs.py
+
 
 class SdLBFGS():
     def __init__(self, func, initial_val, *args,
             max_iterations=1000,
             mem_size=10,
             batch_size=50,
-<<<<<<< Updated upstream:sqn/optimizer/sd_lbfgs.py
             init_step_size=0.1,
             step_size=harmonic_sequence,
             delta=0.1,
             tol=1e-6,
-=======
-            step_size=default_step_size,
-            delta=0.01,
-            tol=1e-4,
->>>>>>> Stashed changes:optimizer/sd_lbfgs.py
             **kwargs):
 
         self._func = func
@@ -69,17 +57,12 @@ class SdLBFGS():
         self._ybars = ShiftList(self._max_mem_size)
         self._rhos = ShiftList(self._max_mem_size)
 
-<<<<<<< Updated upstream:sqn/optimizer/sd_lbfgs.py
         self._iteration_vals = [self._current_val]
         self._iteration_grads = [self._current_grad]
         self._iteration_objvals = [self._current_objval]
 
         self._start_time = time()
         self._iteration_runtimes = [0]
-=======
-        self._iteration_vals = []
-        self._iteration_grad_errs = []
->>>>>>> Stashed changes:optimizer/sd_lbfgs.py
 
         self._result = OptimizeResult()
         self._result['success'] = False
[…]

### 2018-05-07 sqn/optimizer/nonconvex_x6_script.py

@@ -1,10 +1,19 @@
 # -*- coding: utf-8 -*-
+from sys import path
+path.append('../../optimizer/')
+
+from sd_lbfgs import SdLBFGS
+
 from sd_lbfgs import SdLBFGS
 from sd_lbfgs import harmonic_sequence as harmonic_seq
 from sd_lbfgs import sqrt_sequence as sqrt_seq
+
+import algorithms as alg
 import matplotlib.pyplot as plt
 import numpy as np
-from sd_lbfgs import SdLBFGS
+import seaborn as sns
+
+sns.set_context('paper')
 
 
 #%%
@@ -30,20 +39,22 @@ optimizer = SdLBFGS(obj_f, initial_x,
 
 result = optimizer.run()
 
-sdlbfs_x =result['iteration_vals']
-sdlbfs_values =result['iteration_objvals']
-sdlbfs_runtimes =result['iteration_runtimes']
+sdlbfgs_x =result['iteration_vals']
+sdlbfgs_values =result['iteration_objvals']
+sdlbfgs_runtimes =result['iteration_runtimes']
 
 
-print('Solution found by sdLBFGS', sdlbfs_x[-1])
-print('Objective function', no_conv_func2(sdlbfs_x[-1],0))
+print('Solution found by sdLBFGS', sdlbfgs_x[-1])
+print('Objective function', no_conv_func2(sdlbfgs_x[-1],0))
 
-sdlbfs_x =sdlbfs_x[0::int(len(sdlbfs_x)/min(len(sdlbfs_runtimes), points_to_plot))]
-sdlbfs_values =sdlbfs_values[0::int(len(sdlbfs_values)/min(len(sdlbfs_runtimes), points_to_plot))]
-sdlbfs_runtimes =sdlbfs_runtimes[0::int(len(sdlbfs_runtimes)/min(len(sdlbfs_runtimes), points_to_plot))]
+sdlbfgs_x =sdlbfgs_x[0::int(len(sdlbfgs_x)/min(len(sdlbfgs_runtimes), points_to_plot))]
+sdlbfgs_values = sdlbfgs_values[0::int(len(sdlbfgs_values)/min(len(sdlbfgs_runtimes), points_to_plot))]
+sdlbfgs_runtimes =sdlbfgs_runtimes[0::int(len(sdlbfgs_runtimes)/min(len(sdlbfgs_runtimes), points_to_plot))]
+sdlbfgs_length = len(sdlbfgs_values)
 
-'SGD'
[…]

### 2018-05-07 optimizer/sd_lbfgs.py
Merge on Nonconvex function script

@@ -4,11 +4,13 @@ Created on Thu Apr 19 02:07:39 2018
 
 @author: vbigmouse
 """
+
 from numpy import dot, zeros, asarray, sqrt, power
 from math import isnan, isinf
 from scipy.optimize import OptimizeResult
 from time import time
 from pprint import pprint
+
 
 def harmonic_sequence(k, init_step_size):
     # For any nonnegative integer
@@ -17,6 +19,7 @@ def harmonic_sequence(k, init_step_size):
 def sqrt_sequence(k, init_step_size):
     # For any nonnegative integer
     return init_step_size /(sqrt(k+1))
+
 
 def default_step_size(k):
     # For any nonnegative integer
@@ -29,9 +32,10 @@ class SdLBFGS():
             mem_size=10,
             batch_size=50,
             init_step_size=0.1,
-            step_size=harmonic_sequence,
-            delta=0.1,
-            tol=1e-6,
+            step_size=default_step_size,
+            delta=0.01,
+            tol=1e-4,
+
             **kwargs):
 
         self._func = func
@@ -57,12 +61,15 @@ class SdLBFGS():
         self._ybars = ShiftList(self._max_mem_size)
         self._rhos = ShiftList(self._max_mem_size)
 
-        self._iteration_vals = [self._current_val]
-        self._iteration_grads = [self._current_grad]
-        self._iteration_objvals = [self._current_objval]
 
+        self._iteration_vals = []
+        self._iteration_grads = []
+        self._iteration_objvals = []
+
+        self._iteration_grad_errs = []
         self._start_time = time()
-        self._iteration_runtimes = [0]
+        self._iteration_runtimes = []
+
 
         self._result = OptimizeResult()
         self._result['success'] = False
@@ -99,10 +106,9 @@ class SdLBFGS():
 
 
         # stopping criterion, sends back to self.run()
-#        if self._iterations % 10 == 0:
-#            print(f'norm of grad:={dot(self._current_grad.T, self._current_grad)}')
+        self._iteration_grad_errs.append(dot(self._current_grad.T, self._current_grad))
+        if self._iteration_grad_errs[-1]  <= self._tolerance:
 
-        if dot(self._current_grad.T, self._current_grad) <= self._tolerance:
             self._result['success'] = True
 
         # compute search direction
@@ -203,6 +209,8 @@ class SdLBFGS():
         self._result['iteration_grads'] = self._iteration_grads
[…]

### 2018-05-07 optimizer/test.py

@@ -242,286 +242,6 @@ line_bfgs = plt.semilogx(np.array([i for i in range(len(bfgs_runtimes))]),
 plt.legend()
 plt.show()
 
-#%%
-'''non-convex function'''
-
-points_to_plot=1000
-max_iterations = 10000
-
-from hw1_functions import no_conv_func2
-obj_f = lambda x : no_conv_func2(x, order=1)
-
-
-initial_x = 2.
-
-optimizer = SdLBFGS(obj_f, initial_x,
-                    batch_size=1,
-                    delta=0.01,
-                    step_size=sqrt_seq,
-#                    step_size=harmonic_seq,
-                    mem_size=20,
-                    init_step_size=0.1,
-                    max_iterations=10000)
-
-result = optimizer.run()
-
-sdlbfs_x =result['iteration_vals']
-sdlbfs_values =result['iteration_objvals']
-sdlbfs_runtimes =result['iteration_runtimes']
-
-
-print('Solution found by sdLBFGS', sdlbfs_x[-1])
-print('Objective function', no_conv_func2(sdlbfs_x[-1],0))
-
-sdlbfs_x =sdlbfs_x[0::int(len(sdlbfs_x)/min(len(sdlbfs_runtimes), points_to_plot))]
-sdlbfs_values =sdlbfs_values[0::int(len(sdlbfs_values)/min(len(sdlbfs_runtimes), points_to_plot))]
-sdlbfs_runtimes =sdlbfs_runtimes[0::int(len(sdlbfs_runtimes)/min(len(sdlbfs_runtimes), points_to_plot))]
-
-'SGD'
-import algorithms as alg
-
-obj_f = lambda x, order: no_conv_func2(x, order)
-initial_x = 2.
-sgd_x, sgd_values, sgd_runtimes, sgd_xs = \
-    alg.subgradient_descent(obj_f, initial_x, max_iterations, 0.01)
-
-print('Solution found by stochastic subgradient descent', sgd_x)
-print('Objective function', obj_f(sgd_x,0))
-sgd_its = len(sgd_runtimes)
-sgd_values=[obj_f(sgd_xs[i],0) for i in range(0,sgd_its,int(sgd_its/min(sgd_its,points_to_plot)))]
-#%%
-'Ada'
-obj_f = lambda x, order: no_conv_func2(x, order)
[…]

### 2018-05-07 sqn/optimizer/nonconvex_x6_script.py

@@ -20,7 +20,7 @@ sns.set_context('paper')
 '''non-convex function'''
 
 points_to_plot=100
-max_iterations = 10000
+max_iterations = 1000
 
 from hw1_functions import no_conv_func2
 obj_f = lambda x : no_conv_func2(x, order=1)
@@ -35,22 +35,20 @@ optimizer = SdLBFGS(obj_f, initial_x,
 #                    step_size=harmonic_seq,
                     mem_size=20,
                     init_step_size=0.1,
-                    max_iterations=10000)
+                    max_iterations=1000)
 
 result = optimizer.run()
 
 sdlbfgs_x =result['iteration_vals']
 sdlbfgs_values =result['iteration_objvals']
 sdlbfgs_runtimes =result['iteration_runtimes']
+sdlbfs_grads = result['iteration_grads']
 
 
 print('Solution found by sdLBFGS', sdlbfgs_x[-1])
 print('Objective function', no_conv_func2(sdlbfgs_x[-1],0))
 
-sdlbfgs_x =sdlbfgs_x[0::int(len(sdlbfgs_x)/min(len(sdlbfgs_runtimes), points_to_plot))]
-sdlbfgs_values = sdlbfgs_values[0::int(len(sdlbfgs_values)/min(len(sdlbfgs_runtimes), points_to_plot))]
-sdlbfgs_runtimes =sdlbfgs_runtimes[0::int(len(sdlbfgs_runtimes)/min(len(sdlbfgs_runtimes), points_to_plot))]
-sdlbfgs_length = len(sdlbfgs_values)
+sdlbfs_grads = sdlbfs_grads[0::int(sdlbfs_its/min(sdlbfs_its, points_to_plot))]
 
 
 """SGD"""
@@ -58,7 +56,7 @@ sdlbfgs_length = len(sdlbfgs_values)
 
 obj_f = lambda x, order: no_conv_func2(x, order)
 initial_x = 2.
-sgd_x, sgd_values, sgd_runtimes, sgd_xs = \
+sgd_x, sgd_values, sgd_runtimes, sgd_xs, sgd_grads = \
     alg.subgradient_descent(obj_f, initial_x, max_iterations, 0.01)
 
 print('Solution found by stochastic subgradient descent', sgd_x)
[…]

### 2018-05-07 sqn/optimizer/nonconvex_x6_script.py
Discussion section of the paper draft

@@ -11,7 +11,7 @@ from sd_lbfgs import sqrt_sequence as sqrt_seq
 import algorithms as alg
 import matplotlib.pyplot as plt
 import numpy as np
-from sd_lbfgs import SdLBFGS
+import seaborn as sns
 
 sns.set_context('paper')
 
@@ -20,7 +20,7 @@ sns.set_context('paper')
 '''non-convex function'''
 
 points_to_plot=100
-max_iterations = 1000
+max_iterations = 10000
 
 from hw1_functions import no_conv_func2
 obj_f = lambda x : no_conv_func2(x, order=1)
@@ -35,23 +35,22 @@ optimizer = SdLBFGS(obj_f, initial_x,
 #                    step_size=harmonic_seq,
                     mem_size=20,
                     init_step_size=0.1,
-                    max_iterations=1000)
+                    max_iterations=10000)
 
 result = optimizer.run()
 
-sdlbfs_x =result['iteration_vals']
-sdlbfs_values =result['iteration_objvals']
-sdlbfs_runtimes =result['iteration_runtimes']
-sdlbfs_grads = result['iteration_grads']
+sdlbfgs_x =result['iteration_vals']
+sdlbfgs_values =result['iteration_objvals']
+sdlbfgs_runtimes =result['iteration_runtimes']
 
-print('Solution found by sdLBFGS', sdlbfs_x[-1])
-print('Objective function', no_conv_func2(sdlbfs_x[-1],0))
 
-sdlbfs_its = len(sdlbfs_runtimes)
-sdlbfs_x =sdlbfs_x[0::int(sdlbfs_its/min(sdlbfs_its, points_to_plot))]
-sdlbfs_values =sdlbfs_values[0::int(sdlbfs_its/min(sdlbfs_its, points_to_plot))]
-sdlbfs_runtimes =sdlbfs_runtimes[0::int(sdlbfs_its/min(sdlbfs_its, points_to_plot))]
-sdlbfs_grads = sdlbfs_grads[0::int(sdlbfs_its/min(sdlbfs_its, points_to_plot))]
+print('Solution found by sdLBFGS', sdlbfgs_x[-1])
+print('Objective function', no_conv_func2(sdlbfgs_x[-1],0))
+
[…]

### 2018-05-07 optimizer/hw4_functions.py
Add script for SVM loss 1. tanh sigmoid loss(Higges_tanhobj.py) 2. hinge loss(SVM_obj.py)

@@ -1,0 +1,170 @@
+######
+######  This file includes different functions used in HW3
+######
+
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+
+import random
+import math
+
+
+
+def svm_objective_function(w, data, order):
+    labels = data[:,-1]
+    features = data[:,:-1]
+
+    n=len(labels)
+
+    subgradient = np.zeros(w.shape)
+#    result = []
+    i = 0
+#    sample_ind = random.sample(range(0, n), 10000)
+    if order==0:
+#        for w_v in w.T:
+#            if i % 10 == 0:
+        value = 0
+        for i in range(0, n):
+            value += int(max(1 - np.dot(labels[i], np.dot(features[i], w)), 0))
+#        result.append(value)
+        return value
+#        return result
+    elif order==1:
+        for i in range(0, n):
+            if labels[i].item(0) * features[i] * w < 1:
+                subgradient += -np.multiply(labels[i], features[i].T) / n
+            value += max(1 - labels[i] * np.dot(features[i].T, w), 0)
+        return (value, subgradient)
+    else:
+        raise ValueError("The argument \"order\" should be 0 or 1")
+
+def svm_objective_function_stochastic(w, data, order, minibatch_size):
+
+    labels = data[:,-1]
+    features = data[:,:-1]
+    n = len(labels)
+    value = 0
+    subgradient = np.zeros(w.shape)
+    sample_ind = random.sample(range(0, n), minibatch_size)
+
+    if order==0:
+        for i in sample_ind:
+            value += max(1 - labels[i] * features[i] * w , 0)
+        return value
+    elif order==1:
+        for i in sample_ind:
+            if labels[i].item(0) * features[i] * w < 1:
+                subgradient += -np.multiply(labels[i], features[i].T) / minibatch_size
+            value += max(1 - labels[i] * features[i] * w, 0)
+
+
+        return (value, subgradient)
+    else:
+        raise ValueError("The argument \"order\" should be 0 or 1")
+
+################################################################
+
+
+def logrithm_loss(w, data, order, minibatch_size):
+    labels = data[:,-1]
+    features = data[:,:-1]
+    n=len(labels)
+    value = 0
+    subgradient = np.zeros(w.shape)
+
+    sample_ind = random.sample(range(0, n), minibatch_size)
+
+    if order==0:
+        for i in sample_ind:
+            x = features[i].T
[…]

### 2018-05-07 optimizer/sd_lbfgs.py

@@ -35,7 +35,7 @@ class SdLBFGS():
             step_size=default_step_size,
             delta=0.01,
             tol=1e-4,
-
+            damp=True,
             **kwargs):
 
         self._func = func
@@ -48,6 +48,7 @@ class SdLBFGS():
         self._step_size = lambda k:step_size(k, self._init_step_size)
         self._init_step_size = init_step_size
         self._delta = delta
+        self._damp = damp
         self._tolerance = tol
 
         self._iterations = 0
@@ -151,8 +152,8 @@ class SdLBFGS():
                 rho = 0
             else:
                 rho = 1. / sTy
-
-#        theta =1
+        if not self._damp:
+            theta =1
 
         y_bar = theta * y + ((1. - theta) * gamma) * s
 

### 2018-05-07 optimizer/svm_obj.py

@@ -1,0 +1,218 @@
+# -*- coding: utf-8 -*-
+# -*- coding: utf-8 -*-
+from sd_lbfgs import SdLBFGS
+from sd_lbfgs import harmonic_sequence as harmonic_seq
+from sd_lbfgs import sqrt_sequence as sqrt_seq
+import matplotlib.pyplot as plt
+import numpy as np
+
+def norm_z(x):
+    x=(x-x.mean())/x.std()
+    return x
+
+def correct_rate(w, features, labels):
+    cor = 0
+    for i in range(len(features)):
+        if np.dot(w.T,features[i].T) * labels[i] >= 0:
+            cor += 1
+    return cor / len(features)
+
+'''tanh obj test'''
+from hw4_functions import svm_objective_function_stochastic as svm_obj
+
+rawdata = np.loadtxt('HIGGS_subset.csv', delimiter=',')
+labels = np.asmatrix(2 * rawdata[:, 0] -1).T
+features = np.asmatrix(rawdata[:, 1:])
+features = norm_z(features)
+
+data = np.concatenate((features, labels), axis = 1)
+d = features.shape[1]
+
+#%%
+'''General Setting'''
+points_to_plot=100
+max_iterations=10000
+
+
+def const_step(k, initk):
+    return initk
+
+initial_x=np.zeros((d,1))
+
+optimizer = SdLBFGS(lambda x: svm_obj(x, order=1, data=data, minibatch_size=1),
+                    initial_val=initial_x,
+                    batch_size=10,
+                    tol=1e-6,
+                    delta=0.1,
+#                    step_size=harmonic_seq,
+                    step_size=lambda k, initk:const_step(k, initk),
+                    mem_size=20,
+                    init_step_size=0.1,
+                    max_iterations=1000)
+
+result = optimizer.run()
+
+sdlbfs_x =result['iteration_vals']
+sdlbfs_values =result['iteration_objvals']
+sdlbfs_runtimes =result['iteration_runtimes']
+sdlbfs_grads = result['iteration_grads']
+
+
+print('Solution found by sdLBFGS', sdlbfs_x[-1])
[…]

### 2018-05-07 paper/main.tex

@@
 
+We evaluate our implementation of SdLBFGS on a collection of test functions.
+Figure \ref{fig:convex-linear-system} shows the effect of changing the memory
+capacity on a relatively small fully-observable convex differentiable problem.
+Figure \ref{fig:nonconvex-comparison}.
+
+We do not report on results of using the \texttt{SQNOptimizer} to solve
+standard neural network problems in the \texttt{TensorFlow} framework.
+Preliminary results suggest that the present implementation scales
+insufficiently well for use in large scale machine learning problems.
+
 \begin{figure}[h]
@@
   }
-  \label{fig:convex-linear-system}
+  \label{fig:nonconvex-comparison}
 \end{figure}
@@
 
-\begin{enumerate}
-  \item Explaining why the basic project --- implementing a useful practical
-    TensorFlow optimzier --- failed.
-  \item Takeaways from the results.
-  \item Future work.
-\end{enumerate}
+In this work we sought to develop an implementation of SdLBFGS which can be used
+to solve large scale machine learning problems using \texttt{TensorFlow}. By
+this measure, we have reported a negative result. Our preliminary results
+suggest that our implementation, broadly speaking, works satisfactorily on small
+problems, but our attempts at using the \texttt{SQNOptimizer} on problems of the
+size of a typical neural network have been proven infeasible.
+
+Might we say with certainty, then, that the external optimizer approach taken in
+this paper is simply insufficient for solving problems that are typical for the
+\texttt{TensorFlow} framework? Much analysis of our implementation remains
+before we may reject with certainty our approach. The fundamental obstacles we
+encountered were sensitivities to algorithm parameter choices and overall speed
+in computation, both of which yield many possibilities for future work. In the
+former, there may exist choices of step size $\alpha_k$ which perform
+sufficiently admirably for this application, or the trade off frontier between
+memory capacity and computational feasibility may yet become apparent.
+
+In the latter, many directions exist through which the performance of our native
+Python and \texttt{NumPy} implementation might improve significantly. Having
+identified and dealt with potential bottlenecks around the main body of the
+SdLBFGS update, the judicious introduction of performance optimizing tools could
+bring the runtime cost of the implementation into relative feasibility. Such
+tools exist inside the Python ecosystem, from JIT-enhancing through Numba to
+static typing optimizations with Cython. In many cases significant speedups may
+be found by merely modifying the underlying data structures or by adopting
+lazier evaluation techniques.
+
+The central point is that we cannot rule out the
+\texttt{ExternalOptimizerInterface} approach until the exploration of these
+potential improvements has sufficiently resolved the questions raised above.
+Further effort is needed to resolve this partial result.
+
+% \begin{enumerate}
+%   \item Explaining why the basic project --- implementing a useful practical
+%     TensorFlow optimzier --- failed.
+%   \item Takeaways from the results.
+%   \item Future work.
+% \end{enumerate}
 

### 2018-05-07 paper/main.tex
Final commit...?

@@
 \begin{abstract}
-  We do some shit and it doesn't work.
+  We implement a stochastic damped limited memory version of the classical BFGS
+  quasi-Newton minimization algorithm for use in the \texttt{TensorFlow}
+  graphical computation and neural network framework. The implementation is
+  linked to \texttt{TensorFlow} as an \texttt{ExternalOptimizerInterface}
+  object, which circumvents architectural challenges in implementing
+  second-order methods. We provide empirical data to show that the
+  implementation can solve minimization problems in line with its peers. We do
+  not report results on typical problems that are solved using
+  \texttt{TensorFlow} because we find problems of this scale are presently
+  computationally intractible for our implementation.
 \end{abstract}
@@
 
-\begin{enumerate}
-  \item Review main results of the Wang, Ma, Goldfarb, Liu paper
-    \cite{sdlbfgs}
-  \item State the SdLBFGS algortihm.
-  \item Mention some details about how the algorithm translates to code.
-\end{enumerate}
+% \begin{enumerate}
+%   \item Review main results of the Wang, Ma, Goldfarb, Liu paper
+%     \cite{sdlbfgs}
+%   \item State the SdLBFGS algortihm.
+%   \item Mention some details about how the algorithm translates to code.
+% \end{enumerate}
 
@@
 capacity on a relatively small fully-observable convex differentiable problem.
-Figure \ref{fig:nonconvex-comparison}.
+Figure \ref{fig:nonconvex-comparison}. Additionally, we compare the performance
+of our implementation of SdLBFGS with other common minimizers on a
+dataset\footnote{The data was provided by Professor Orabona.} for a Higgs boson
+classification problem. Empirical results in Figure \ref{fig:svm-hinge} are
+interesting simply by virtue of lacking any theoretical justification for
+convergence, as the loss function is not $C^1$ with Lipschitz gradient. The
+results when the loss function is in $C^1$ with Lipschitz gradient, as in Figure
+\ref{fig:svm-tan}, however, do conform to the theoretical guarantees.
 
@@
   \label{fig:convex-linear-system}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\textwidth]{../plots/svm_hinge_loss.eps}
+  \caption{
+    SdLBFGS performance on the Higgs boson dataset using an SVM classifier with
+    a hinge loss function, compared with other minimization algorithms. Though
+    no theoretical guarantees are made about the performance of SdBFGS, we
+    observe that it behaves quite well with respect to its peers. The outlier
+    performance is the standard L-BFGS algorithm, which seems to behave poorly
+    on this problem.
+  }
+  \label{fig:svm-hinge}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\textwidth]{../plots/svm_tan_regularized_loss.eps}
+  \caption{
+    SdLBFGS performance on the Higgs boson dataset using an SVM classifier with
+    a smooth regularized function, compared with other minimization algorithms.
+    Unlike in Figure \ref{fig:svm-hinge}, we do have theoretical convergence
+    guarantees for this $f(x)$. We also
+    observe that it behaves quite well with respect to its peers. The outlier
+    performance is again the standard L-BFGS algorithm, which seems to behave poorly
+    on this problem.
+  }
+  \label{fig:svm-tan}
 \end{figure}

### 2018-05-07 optimizer/Higges_tanhobj.py

@@ -4,6 +4,9 @@ from sd_lbfgs import harmonic_sequence as harmonic_seq
 from sd_lbfgs import sqrt_sequence as sqrt_seq
 import matplotlib.pyplot as plt
 import numpy as np
+import seaborn as sns
+
+sns.set_context('paper')
 
 def norm_z(x):
     x=(x-x.mean())/x.std()
@@ -151,70 +154,108 @@ lbfgs_grads = lbfgs_grads[0::int(lbfgs_its/min(lbfgs_its, points_to_plot))]
 
 #%%
 '''num iterations to gradient norm'''
-plt.figure(figsize=(14,8))
+
+fig, axes = plt.subplots(1,2, figsize=(12,4))
+
+plot_settings = {
+        'linewidth': 2,
+        'dashes': [1,1]
+        }
+
 
 
 sdlbfgs_X = np.array([i for i in range(0, len(sdlbfs_values))])
 sdlbfgs_Y = np.linalg.norm(np.array(sdlbfs_grads), axis=1).reshape(sdlbfgs_X.shape)
-line_sdlbfgs, = plt.semilogx( sdlbfgs_X, sdlbfgs_Y, linewidth=2, color='k', dashes = [1, 1],
-                         marker='.', label='SdLBFGS')
+line_sdlbfgs, = axes[0].semilogx(
+        sdlbfgs_X, 
+        sdlbfgs_Y, 
+        marker='v', 
+        label='SdLBFGS', 
+        **plot_settings)
 
 sgd_X = np.array([i for i in range(len(sgd_values))])
 sgd_Y = np.linalg.norm(np.array(sgd_grads), axis=1).reshape(sgd_X.shape)
-line_sgd, = plt.semilogx(sgd_X,sgd_Y, linewidth=2, color='r', dashes = [1, 1],
-                         marker='o', label='SGD')
+line_sgd, = axes[0].semilogx(
+        sgd_X,
+        sgd_Y, 
+        marker='o', 
+        label='SGD',
+        **plot_settings)
 
 ada_X = np.array([i for i in range(len(ada_values))])
 ada_Y = np.linalg.norm(np.array(ada_grads),axis=1).reshape(ada_X.shape)
-line_ada, = plt.semilogx(ada_X, ada_Y, linewidth=2, color='b', dashes = [4, 1],
-                        marker='x', label='AdaGrad')
+line_ada, = axes[0].semilogx(
+        ada_X, 
+        ada_Y, 
[…]

### 2018-05-07 optimizer/svm_obj.py

@@ -5,6 +5,9 @@ from sd_lbfgs import harmonic_sequence as harmonic_seq
 from sd_lbfgs import sqrt_sequence as sqrt_seq
 import matplotlib.pyplot as plt
 import numpy as np
+import seaborn as sns
+
+sns.set_context('paper')
 
 def norm_z(x):
     x=(x-x.mean())/x.std()
@@ -144,70 +147,107 @@ lbfgs_runtimes =lbfgs_runtimes[0::int(lbfgs_its/min(lbfgs_its, points_to_plot))]
 lbfgs_grads = lbfgs_grads[0::int(lbfgs_its/min(lbfgs_its, points_to_plot))]
 #%%
 '''num iterations to gradient norm'''
-plt.figure(figsize=(14,8))
+fig, axes = plt.subplots(1,2, figsize=(12,4))
+
+plot_settings = {
+        'linewidth': 2,
+        'dashes': [1,1]
+        }
 
 
 sdlbfgs_X = np.array([i for i in range(0, len(sdlbfs_values))])
 sdlbfgs_Y = np.linalg.norm(np.array(sdlbfs_grads), axis=1).reshape(sdlbfgs_X.shape)
-line_sdlbfgs, = plt.semilogx( sdlbfgs_X, sdlbfgs_Y, linewidth=2, color='k', dashes = [1, 1],
-                         marker='.', label='SdLBFGS')
+line_sdlbfgs, = axes[0].semilogx(
+        sdlbfgs_X, 
+        sdlbfgs_Y, 
+        marker='v', 
+        label='SdLBFGS',
+        **plot_settings)
 
 sgd_X = np.array([i for i in range(len(sgd_values))])
 sgd_Y = np.linalg.norm(np.array(sgd_grads), axis=1).reshape(sgd_X.shape)
-line_sgd, = plt.semilogx(sgd_X,sgd_Y, linewidth=2, color='r', dashes = [1, 1],
-                         marker='o', label='SGD')
+line_sgd, = axes[0].semilogx(
+        sgd_X,
+        sgd_Y, 
+        marker='o', 
+        label='SGD',
+        **plot_settings)
 
 ada_X = np.array([i for i in range(len(ada_values))])
 ada_Y = np.linalg.norm(np.array(ada_grads),axis=1).reshape(ada_X.shape)
-line_ada, = plt.semilogx(ada_X, ada_Y, linewidth=2, color='b', dashes = [4, 1],
[…]
